# Getting started

<div class="callout-caution">
The functions presented below constitute the core wrapper functions utilized throughout the analysis, including three functions originally from the @imbens2024 tutorial, that were modified to suit specific analytical purposes (highlighted by ***), as well as others that were integrated without modification. For the full collection of functions by @imbens2024, please consult their official source code available on their GitHub [repository](https://github.com/lkrsl/Application-of-Outcome-Weights-Framework-for-DML-to-Lalonde-Study).
</div>

## Installation

Several R packages are required for the data analysis and visualization. The code below checks for all required packages and installs those that are missing. Once installation is complete, the packages must be loaded.  

<div class="callout-note">
Since the analysis sources functions directly from the @imbens2024 tutorial, the required packages are automatically installed and loaded as part of that process. Therefore, it is only necessary to install and load any additional packages as well as add custom functions that are required for the analysis.
</div>

**Packages**: "cobalt", "data.table", "dplyr", "ebal", "ggplot2", "gridExtra", "highr", "highs" , "MatchIt", "optmatch", "optweight", "quickmatch", "readr", "rgenoud", "tidyr", "tidyverse", "ppcor", "WeightIt"
```{r, message=FALSE, warning=FALSE}
# required packages
packages <- c("cobalt", "data.table", "dplyr", "ebal", "ggplot2", "gridExtra", "highr", "highs", 
              "kableExtra", "MatchIt", "optmatch", "optweight", "quickmatch", 
              "readr", "rgenoud", "tidyr", "tidyverse", "ppcor", "WeightIt"
)

# install packages
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}

install_all(packages)

# load packages
library(cobalt)
library(data.table)
library(dplyr)
library(ebal)
library(ggplot2)
library(gridExtra)
library(highr)
library(highs)
library(kableExtra)
library(MatchIt)
library(optmatch)
library(optweight)
library(quickmatch)
library(readr)
library(rgenoud)
library(tidyr)
library(tidyverse)
library(ppcor)
library(WeightIt)
```

## Wrapper functions

Next, the wrapper functions designed to manage the datasets, address overlap and covariate balance, estimate treatment effects and analyze outcome weights are presented.

<div class="callout-note">
To use the wrapper functions, the simplest method is to source the R script with the following code.
</div>
```{r, message=FALSE, warning=FALSE}
source("tutorial/functions.R")
```

If RStudio is used, the functions should now appear in the environment section.

### Overview

| Function                                            | Description                                                                                                                                                                     |                 Code                  |
|------------------|------------------------------------|:----------------:|
| `inspect_data()` | Inspects datasets.    | [More](#inspec) |
| `truncate_weights_fixed()`, `truncate_weights_percentile()`, `check_weights()`, `truncate_weights_adaptive()`, `ps_estimate()`, `ps_trim()`, `common_range_trim()`, `crump_trim()`, `stuermer_trim ()`, `walker_trim()`, `trim_attach_weights()` | Implements methods to improve covariate balancing and overlap.   | [More](#impl-cov-bal-overl-mthds)  |
| `get_smd_stats()`, `compute_abs_smd_matchit()`, `compute_ess_matchit()`, `plot_matching_balance()`,
`compute_abs_smd_weight()`, `compute_ess_weight()`, `plot_weighting_balance()`, `compute_abs_smd_trunc()`, `compute_ess_trunc()`, `plot_trunc_balance()`, `compute_abs_smd_trim()`, `compute_ess_trim()`, `plot_trim_overlap()`, `compute_abs_smd_comb()`, `compute_ess_comb()`, `plot_comb_overlap()`, `save_comb_hist()`, `combine_results()`, `combine_results_plus()`, `save_csv()`, `assess_methods()`, `get_top_methods()`, `create_top5_datasets()`, `save_top5_datasets()` | Evaluates performance of previous methods through comparative metrics (absolute standardized mean differences (SMD) and effective sample size (ESS)).    | [More](#eval-mthds) |
| `estimate_all()`, `plot_coef()`, `save_att_panels()`, `create_matrix_results()`, `eval_att()`, `plot_catt_panels()`, `save_catt_panels()`, `save_main_catt_panels()`, `save_plus_catt_panels()`, `eval_catt()`, `plot_qte_top()`, `save_qtet()`, `save_qte_top()` | Computes and visualizes the Average Treatment Effect on the Treated (ATT) using a number of estimators.    | [More](#est-and-plot-att-catt-qtet) |
| `get_res_att()`, `derive_ow()`, `plot_ow()`, `eval_ow()`, `save_ow()` | Analyses outcome weights of datasets.    | [More](#ow-ana) |
| `check_filter_datasets()`, `save_sensitivity_plots()` | Validates and filters datasets.    | [More](#sens-ana) |

<a name="inspec"></a>

### Inspection

`inspect_data()` provides a structured summary of one or multiple datasets by returning a single data frame listing each dataset’s name, number of observations (rows), number of treated units, number of control units, number of variables (columns) and a concatenated string of variable names.
```{r}
inspect_data <- function(data, treat = "treat") {
  if (is.data.frame(data)) {
    data <- list(dataset = data)
  }
  data.frame(
    dataset = names(data),
    num_obs = sapply(data, nrow),
    num_treated = sapply(data, function(df) sum(df[[treat]] == 1, na.rm = TRUE)),
    num_controls = sapply(data, function(df) sum(df[[treat]] == 0, na.rm = TRUE)),
    num_vars = sapply(data, ncol),
    name_vars = sapply(data, function(df) paste(names(df), collapse = ", ")),
    row.names = NULL
  )
}
```

**Argument**

- `data_list`: A data frame or a list of data frames containing structured data.

<br>

<a name="impl-cov-bal-overl-mthds"></a>

### Improving covariate balance and overlap
#### Matching
NA

#### Weighting
NA

#### Truncation

`truncate_weights_fixed()` caps weights at a fixed maximum range (@Larson_2023, @Xiao_2013, Matsouaka_2023).
```{r}
truncate_weights_fixed <- function(data, weight_col, lower = 0.025, upper = 0.975) {
  data[[weight_col]] <- pmax(data[[weight_col]], lower)
  data[[weight_col]] <- pmin(data[[weight_col]], upper)
  return(data)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `weight_col`: The column name of the weights in the dataset as string.
- `lower`: Numeric value specifying the minimum allowed weight. Weights below this value are set to this lower bound.
- `upper`: Numeric value specifying the maximum allowed weight. Weights above this value are set to this upper bound.

`truncate_weights_percentile()` caps weights at specified percentile cutpoints (@Larson_2023, @Cole_Hernan_2008, @Matsouaka_2023).
```{r}
truncate_weights_percentile <- function(data, weight_col, lower = 0.05, upper = 0.95) {
  quantiles <- quantile(data[[weight_col]], probs = c(lower, upper), na.rm = TRUE)
  data[[weight_col]] <- pmax(data[[weight_col]], quantiles[1])
  data[[weight_col]] <- pmin(data[[weight_col]], quantiles[2])
  return(data)
} 
```

**Arguments**

- `data`: The data frame containing structured data.
- `weight_col`: The column name of the weights in the dataset as string.
- `lower`: Numeric value between 0 and 1 specifying the lower quantile cutpoint. Weights below this percentile are set to the value at this quantile.
- `upper`: Numeric value between 0 and 1 specifying the upper quantile cutpoint. Weights above this percentile are set to the value at this quantile.

`truncate_weights_adaptive()` adaptively selects optimal lower and upper quantile cutpoints for truncating a precomputed weights by performing cross-validation over quantile grids (@Schwab_2017).
```{r}
truncate_weights_adaptive_weights <- function(
    data, weight_col, folds = 5,
    lower_grid = seq(0.01, 0.10, by = 0.01),
    upper_grid = seq(0.90, 0.99, by = 0.01),
    loss_fn = function(w) var(w, na.rm=TRUE)
) {
  n <- nrow(data)
  fold_ids <- sample(rep(1:folds, length.out = n))
  loss_mat <- matrix(NA, nrow = length(lower_grid), ncol = length(upper_grid))
  for (i in seq_along(lower_grid)) {
    for (j in seq_along(upper_grid)) {
      if (lower_grid[i] >= upper_grid[j]) next
      cv_losses <- c()
      for (fold in 1:folds) {
        valid_idx <- which(fold_ids == fold)
        valid_weights <- data[[weight_col]][valid_idx]
        lower_q <- quantile(valid_weights, probs = lower_grid[i], na.rm = TRUE)
        upper_q <- quantile(valid_weights, probs = upper_grid[j], na.rm = TRUE)
        w_trunc <- pmax(pmin(valid_weights, upper_q), lower_q)
        cv_losses <- c(cv_losses, loss_fn(w_trunc))
      }
      loss_mat[i,j] <- mean(cv_losses)
    }
  }
  idx <- which(loss_mat == min(loss_mat, na.rm = TRUE), arr.ind = TRUE)
  best_lower <- lower_grid[idx[1]]
  best_upper <- upper_grid[idx[2]]
  quantiles <- quantile(data[[weight_col]], probs = c(best_lower, best_upper), na.rm = TRUE)
  data[[weight_col]] <- pmax(pmin(data[[weight_col]], quantiles[2]), quantiles[1])
  return(data)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `weight_col`: The column name of the weights in the dataset as string.
- `folds`: Integer specifying the number of cross-validation folds to use when searching for the optimal quantiles.
- `lower_grid`: Numeric vector of candidate lower quantile cutpoints to search over (between 0 and 1).
- `upper_grid`: Numeric vector of candidate upper quantile cutpoints to search over (between 0 and 1).

#### Trimming 
`ps_estimate()` reestimates the propensity scores by fitting a probability forest with the treatment indicator as the target and the specified covariates as predictors.
```{r}
ps_estimate <- function(data, Y, treat, covar, num.trees = 4000, seed = 42) {
  set.seed(seed)
  data$ps_estimate <- probability_forest(
    X = data[, covar],
    Y = as.factor(data[, treat]),
    seed = seed, num.trees = num.trees
  )$predictions[, 2]
  return(data)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `Y`: String specifying the name of the outcome variable.
- `treat`: The column name of the (binary) treatment indicator as string (default is `treat`). 
- `covar`: A vector of covariate names.
- `num.trees`:Integer specifying the number of trees to grow in the random forest model for propensity score estimation (default is 4000).
- `seed`:  Integer to ensure reproducibility (default is 42).

`ps_trim()` trims the dataset based on a specified propensity score threshold, removing observations whose propensity scores exceed the given cutoff value.*** 
```{r}
ps_trim <- function(data, ps = "ps_assoverlap", threshold = 0.9) { 
  sub <- data[which(data[, ps] < threshold), ]
  return(sub)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `ps`: The name of the propensity score variable as string.
- `threshold`: The upper limit for propensity score inclusion as numeric value.

`common_range_trim()` trims the dataset to observations with propensity scores within the common support range for treated and control groups. It takes as lower cutpoint the lowest observed propensity score in the treated group, and as upper cutpoint the highest observed propensity score in the control group.
```{r}
common_range_trim <- function(data, ps = "ps_assoverlap", treat = "treat") {
  lower_cut <- min(data[[ps]][data[[treat]] == 1], na.rm = TRUE)
  upper_cut <- max(data[[ps]][data[[treat]] == 0], na.rm = TRUE)
  sub <- data[data[[ps]] >= lower_cut & data[[ps]] <= upper_cut, ]
  return(sub)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `ps`: The column name of propensity scores as string (default is `ps_assoverlap`). 
- `treat`: The column name of the (binary) treatment indicator as string (default is `treat`). 

`crump_trim()` trims the dataset to observations with propensity scores outside specified lower and upper bounds (default 0.1 and 0.9). Only those observations are kept that are within the specified interval (default 0.1 <= PS <= 0.9).
```{r}
crump_trim <- function(data, ps = "ps_assoverlap", lower = 0.1, upper = 0.9) {
  sub <- data[data[[ps]] >= lower & data[[ps]] <= upper, ]
  return(sub)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `ps`: The column name of propensity scores as string (default is `ps_assoverlap`). 
- `lower`: The lower bound for inclusion as numeric value (default is 0.1).
- `upper`: The upper bound for inclusion as numeric value (default is 0.9).

`stuermer_trim ()` trims the dataset to observations based on propensity score quantiles separately for treated and control group.
```{r}
stuermer_trim <- function(data, treat = "treat", ps = "ps_assoverlap", 
                          lower_percentile = 0.05, upper_percentile = 0.95) {
  treated_ps   <- data[[ps]][data[[treat]] == 1]
  untreated_ps <- data[[ps]][data[[treat]] == 0]
  lower_cutoff <- quantile(treated_ps, probs = lower_percentile, na.rm = TRUE)
  upper_cutoff <- quantile(untreated_ps, probs = upper_percentile, na.rm = TRUE)
  sub <- data[data[[ps]] >= lower_cutoff & data[[ps]] <= upper_cutoff, ]
  return(sub)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `treat`: The column name of the (binary) treatment indicator as string (default is `treat`). 
- `ps`: The column name of propensity scores as string (default is `ps_assoverlap`).
- `lower_percentile`: Numeric value between 0 and 1 specifying the lower quantile cutoff for propensity scores in the treated group (default is 0.05, i.e., 5th percentile).
- `upper_percentile`: Numeric value between 0 and 1 specifying the upper quantile cutoff for propensity scores in the untreated group (default is 0.95, i.e., 95th percentile).

`walker_trim()` trims the dataset to observations based on preference scores that adjust for treatment prevalence using logit transformations.
```{r}
walker_trim <- function(data, treat = "treat", ps = "ps_assoverlap", 
                        lower_cutoff = 0.3, upper_cutoff = 0.7) {
  treat_prevalence  <- mean(data[[treat]], na.rm = TRUE)
  logit_ps          <- log(data[[ps]] / (1 - data[[ps]]))
  logit_prevalence  <- log(treat_prevalence / (1 - treat_prevalence))
  preference_score  <- 1 / (1 + exp(-(logit_ps - logit_prevalence)))
  sub <- data[preference_score >= lower_cutoff & preference_score <= upper_cutoff, ]
  return(sub)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `treat`: The column name of the (binary) treatment indicator as string (default is `treat`). 
- `ps`: The column name of propensity scores as string (default is `ps_assoverlap`).
- `lower_cutoff`: Numeric value between 0 and 1 specifying the lower cutoff for the preference score (default is 0.3).
- `upper_cutoff`: Numeric value between 0 and 1 specifying the upper cutoff for the preference score (default is 0.7).

#### Combined methods
`attach_weights()` merges trimmed data with original weights, preserving corresponding weight values after trimming.
```{r}
attach_weights <- function(trimmed_list, model, weight_type = c("ipw_weight", "opt_weight", "cbps_weight", "ebal_weight"), estimand = "ATT") {
  weight_type <- match.arg(weight_type)
  weight_fun <- function(data) {
    if (weight_type == "ipw_weight") {
      w.obj <- WeightIt::weightit(model, data = data, estimand = estimand, method = "glm")
      data$ipw_weight <- w.obj$weights
    } else if (weight_type == "opt_weight") {
      w.obj <- optweight::optweight(model, data = data, estimand = estimand)
      data$opt_weight <- w.obj$weights
    } else if (weight_type == "cbps_weight") {
      w.obj <- WeightIt::weightit(model, data = data, estimand = estimand, method = "cbps")
      data$cbps_weight <- w.obj$weights
    } else if (weight_type == "ebal_weight") {
      w.obj <- WeightIt::weightit(model, data = data, estimand = estimand, method = "ebal")
      data$ebal_weight <- w.obj$weights
    }
    return(data)
  }
  weighted_list <- lapply(trimmed_list, weight_fun)
  return(weighted_list)
}
```

**Arguments**

- `trimmed_list`: A list of data frames containing the subsetted dataset after trimming procedures.
- `model`: A model formula specifying the treatment and covariates to be used for calculating weights.
- `weight_type`: A character string specifying the type of weighting method to apply when attaching weights to the trimmed datasets.

<br>

<a name="eval-mthds"></a>

### Reassessing methods
#### Matching

`get_smd_stats()` extracts mean and maximum absolute standardized mean differences (SMDs) from a balance object to assess covariate balance.
```{r}
get_smd_stats <- function(match_object) {
  bal <- bal.tab(match_object, stats = "mean.diffs", un = TRUE, s.d.denom = "treated")
  smds <- bal$Balance$Diff.Adj
  smds <- smds[!(rownames(bal$Balance) %in% c("distance"))]
  mean_smd <- mean(abs(smds), na.rm = TRUE)
  max_smd <- max(abs(smds), na.rm = TRUE)
  return(c(Mean_Abs_SMD = mean_smd, Max_Abs_SMD = max_smd))
}
```

**Argument**

- `match_object`: The balance / matching object, containing information on covariate balance.

`compute_abs_smd_matchit()` computes absolute SMDs for a list of MatchIt objects and returns summary statistics.
```{r}
compute_abs_smd_matchit <- function(match_list) {
  smd_list <- lapply(match_list, get_smd_stats)       
  smd_mat <- do.call(rbind, smd_list)                 
  smd_df <- data.frame(
    Method = names(match_list),
    Mean_Abs_SMD = smd_mat[, "Mean_Abs_SMD"],
    Max_Abs_SMD  = smd_mat[, "Max_Abs_SMD"],
    row.names = NULL
  )
  return(smd_df)
} 
```

**Argument**

- `match_list`: A list of MatchIt objects.

`compute_ess_matchit()` extracts and formats effective sample size (ESS) information from the balance summary of a MatchIt object.
```{r}
compute_ess_matchit <- function(bal_tab_object) {
  samples <- bal_tab_object$Observations
  df <- as.data.frame(samples)
  df$Method <- rownames(samples)
  df <- df[, c("Method","Control", "Treated")]
  rownames(df) <- NULL
  df
}
```

`plot_matching_balance()` generates love plots for a list of MatchIt objects, visualizing covariate balance before and after matching.
```{r}
plot_matching_balance <- function(matchit_objects, threshold = 0.1, title = NULL) {
  if (!is.list(matchit_objects)) {
    matchit_objects <- list(single = matchit_objects)
  }
  plot_list <- list()
  for (name in names(matchit_objects)) {
    matchit_object <- matchit_objects[[name]]
    bal <- cobalt::bal.tab(matchit_object, un = TRUE)
    smd_df <- data.frame(
      Variable = rownames(bal$Balance),
      Pre = bal$Balance[,"Diff.Un"],
      Post = bal$Balance[,"Diff.Adj"]
    )
    # remove propensity score or non-covariate rows if present
    smd_df <- smd_df[!grepl("^distance$|^prop.score$|distance|prop.score", smd_df$Variable), ]
    # convert to long format
    smd_long <- pivot_longer(
      smd_df,
      cols = c("Pre", "Post"),
      names_to = "Matching",
      values_to = "Std_Diff"
    )
    smd_long$Matching <- factor(
      smd_long$Matching, 
      levels = c("Pre", "Post"), 
      labels = c("Pre-Matching", "Post-Matching")
    )
    # build titles
    plot_title <- if (length(matchit_objects) > 1) paste(title, "-", name) else title
    # create ggplot 
    p <- ggplot(smd_long, aes(x = Variable, y = Std_Diff, color = Matching)) +
      geom_point(size = 3) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
      geom_hline(yintercept = threshold, linetype = "dashed", color = "red") +
      geom_hline(yintercept = -threshold, linetype = "dashed", color = "red") +
      coord_flip() +
      labs(title = plot_title, x = "Covariates", y = "Standardized Mean Differences") +
      theme_minimal() +
      theme(panel.border = element_rect(color = "black", fill = NA, size = 1) ,
            plot.title = element_text(face = "plain", size = 11, hjust = 0.5)) +
      scale_color_manual(values = c("Pre-Matching" = "#00CFC1", "Post-Matching" = "#FC766A"))
    plot_list[[name]] <- p
  }
  
  if (length(plot_list) == 1) {
    return(plot_list[[1]])
  } else {
    return(plot_list)
  }
}
```

**Argument**

- `matchit_objects`: A single MatchIt object or a named list of MatchIt objects, each containing matched data produced by the MatchIt package for different matching specifications.
- `threshold`: A numeric value specifying the acceptable limit for standardized mean differences (default is 0.1).
- `title`: An optional character string for the main title of the plot or, if plotting multiple objects, a prefix for individual plot titles.

#### Weighting

`compute_abs_smd_weight()` calculates absolute SMDs for datasets weighted by specified weight vectors.
```{r}
compute_abs_smd_weight <- function(data, treat, covar, weights_list) {
  smd_list <- lapply(names(weights_list), function(method) {
    bal_obj <- cobalt::bal.tab(
      as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
      data = data,
      weights = weights_list[[method]],
      un = TRUE,
      s.d.denom = "treated"
    )
    smd_df <- as.data.frame(bal_obj$Balance)
    smd_vals <- abs(smd_df$Diff.Adj)
    mean_smd <- mean(smd_vals, na.rm = TRUE)
    max_smd  <- max(smd_vals, na.rm = TRUE)
    return(data.frame(
      Method = method,
      Mean_Abs_SMD = mean_smd,
      Max_Abs_SMD  = max_smd
    ))
  })
  do.call(rbind, smd_list)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `weights_list`: A list of weight vectors. 

`compute_ess_weight()` extracts ESS associated with different weighting schemes.
```{r}
compute_ess_weight <- function(data, treat, covar, weights_list) {
  ess_list <- lapply(names(weights_list), function(method) {
    bal_obj <- cobalt::bal.tab(
      as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
      data = data,
      weights = weights_list[[method]],
      un = FALSE
    )
    samples <- bal_obj$Observations
    if ("Adjusted" %in% rownames(samples)) {
      df <- samples["Adjusted", c("Control", "Treated"), drop = FALSE]
    } else {
      df <- samples[1, c("Control", "Treated"), drop = FALSE]
    }
    df <- cbind(Method = method, df)
    rownames(df) <- NULL
    return(df)
  })
  do.call(rbind, ess_list)
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `weights_list`: A list of weight vectors. 

`plot_weighting_methods()` generates love plots, visualizing covariate balance under different weighting methods.
```{r}
plot_weighting_balance <- function(data, treat, covar, weight_list, title = NULL) {
  plot_list <- list()
  for (wname in names(weight_list)) {
    invisible(capture.output({
    bal <- cobalt::bal.tab(
      as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
      data = data,
      weights = weight_list[[wname]],
      un = TRUE,
      s.d.denom = "treated"
    )
    smd_df <- data.frame(
      Variable = rownames(bal$Balance),
      Pre  = bal$Balance[,"Diff.Un"],
      Post = bal$Balance[,"Diff.Adj"]
    )
    smd_df <- smd_df[!grepl("^distance$|^prop.score$|distance|prop.score", smd_df$Variable), ]
    smd_long <- pivot_longer(
      smd_df,
      cols = c("Pre", "Post"),
      names_to = "Weighting",
      values_to = "Std_Diff"
    )
    smd_long$Weighting <- factor(
      smd_long$Weighting,
      levels = c("Pre", "Post"),
      labels = c("Pre-Weighting", "Post-Weighting")
    )
    plot_title <- if (length(weight_list) > 1) paste(title, "-", wname) else title
    p <- ggplot(smd_long, aes(x = Variable, y = Std_Diff, color = Weighting)) +
      geom_point(size = 3) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
      geom_hline(yintercept = 0.1, linetype = "dashed", color = "red") +
      geom_hline(yintercept = -0.1, linetype = "dashed", color = "red") +
      coord_flip() +
      labs(title = plot_title, x = "Covariates", y = "Standardized Mean Differences") +
      theme_minimal() +
      theme(panel.border = element_rect(color = "black", fill = NA, size = 1) ,
            plot.title = element_text(face = "plain", size = 11, hjust = 0.5)) +
      scale_color_manual(values = c("Pre-Weighting" = "#00CFC1", "Post-Weighting" = "#FC766A"))
    plot_list[[wname]] <- p
   }))
  }  
  if (length(plot_list) == 1) {
    return(plot_list[[1]])
  } else {
    return(plot_list)
  }
}
```

**Arguments**

- `data`: The data frame containing structured data.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `weight_list`: A list of weight vectors. 
- `dataset_name`: The dataset name for plot titles.

#### Truncation

`compute_abs_smd_trunc()` calculates absolute SMDs for truncated weight datasets across different truncation methods and weighting variables.
```{r}
compute_abs_smd_trunc <- function(trunc_list, treat, covar, weight_cols) {
  all_smd <- list()
  for(trunc_name in names(trunc_list)) {
    dataset <- trunc_list[[trunc_name]]
    smd_list <- lapply(weight_cols, function(wcol) {
      if (wcol %in% names(dataset)) {
        bal_obj <- cobalt::bal.tab(
          as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
          data = dataset,
          weights = dataset[[wcol]],
          un = TRUE,
          s.d.denom = "treated"
        )
        smd_df <- as.data.frame(bal_obj$Balance)
        smd_vals <- abs(smd_df$Diff.Adj)
        data.frame(
          Method = paste(trunc_name, wcol, sep = "_"),
          Mean_Abs_SMD = mean(smd_vals, na.rm = TRUE),
          Max_Abs_SMD = max(smd_vals, na.rm = TRUE)
        )
      } else {
        NULL
      }
    })
    all_smd[[trunc_name]] <- do.call(rbind, smd_list)
  }
  smd_summary <- do.call(rbind, all_smd)
  rownames(smd_summary) <- NULL  
  return(smd_summary)
}
```

**Arguments**

- `trunc_list`: A list of data frames, each containing a dataset with truncated weights.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `weight_cols`: The name of the columns holding the weights.

`compute_ess_trunc()` computes ESS values for truncated weight datasets considering different methods and weights.
```{r}
compute_ess_trunc <- function(trunc_list, treat, covar, weight_cols) {
  all_ess <- list()
  for(trunc_name in names(trunc_list)) {
    dataset <- trunc_list[[trunc_name]]
    ess_list <- lapply(weight_cols, function(wcol) {
      if (wcol %in% names(dataset)) {
        bal_obj <- cobalt::bal.tab(
          as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
          data = dataset,
          weights = dataset[[wcol]],
          un = FALSE,
          s.d.denom = "treated"
        )
        samples <- bal_obj$Observations
        if ("Adjusted" %in% rownames(samples)) {
          df <- data.frame(
            Method = paste(trunc_name, wcol, sep = "_"),
            Control = samples["Adjusted", "Control"],
            Treated = samples["Adjusted", "Treated"]
          )
        } else {
          df <- data.frame(
            Method = paste(trunc_name, wcol, sep = "_"),
            Control = samples[1, "Control"],
            Treated = samples[1, "Treated"]
          )
        }
        df
      } else {
        NULL
      }
    })
    all_ess[[trunc_name]] <- do.call(rbind, ess_list)
  }
  ess_summary <- do.call(rbind, all_ess)
  rownames(ess_summary) <- NULL   
  return(ess_summary)
}
```

**Arguments**

- `trunc_list`: A list of data frames, each containing a dataset with truncated weights.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `weight_cols`: The name of the columns holding the weights.

`plot_trunc_methods()` produces love plots for various truncation methods and weight types, visualizing covariate balance before and after truncation.
```{r}
plot_trunc_balance <- function(trunc_list, treat, covar, weight_cols, dataset_name = NULL, threshold = 0.1) {
  plot_list <- list()
  for (trunc_name in names(trunc_list)) {
    dataset <- trunc_list[[trunc_name]]
    for (wcol in weight_cols) 
      invisible(capture.output({
      if (wcol %in% names(dataset)) {
        bal_obj <- cobalt::bal.tab(
          as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
          data = dataset,
          weights = dataset[[wcol]],
          un = TRUE,
          s.d.denom = "treated"
        )
        smd_df <- data.frame(
          Variable = rownames(bal_obj$Balance),
          Pre = bal_obj$Balance[,"Diff.Un"],
          Post = bal_obj$Balance[,"Diff.Adj"]
        )
        smd_df <- smd_df[!grepl("^distance$|^prop.score$|distance|prop.score", smd_df$Variable), ]
        smd_long <- pivot_longer(
          smd_df,
          cols = c("Pre", "Post"),
          names_to = "Truncation",
          values_to = "Std_Diff"
        )
        smd_long$Truncation <- factor(
          smd_long$Truncation,
          levels = c("Pre", "Post"),
          labels = c("Pre-Truncation", "Post-Truncation")
        )
        plot_title <- paste0(
          if (!is.null(dataset_name)) paste0(dataset_name, " - ") else "",
          trunc_name, "_", wcol, " truncation"
        )
        p <- ggplot(smd_long, aes(x = Variable, y = Std_Diff, color = Truncation)) +
          geom_point(size = 3) +
          geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
          geom_hline(yintercept = threshold, linetype = "dashed", color = "red") +
          geom_hline(yintercept = -threshold, linetype = "dashed", color = "red") +
          coord_flip() +
          labs(title = plot_title, x = "Covariates", y = "Standardized Mean Differences") +
          theme_minimal() +
          theme(panel.border = element_rect(color = "black", fill = NA, size = 1),
                plot.title = element_text(face = "plain", size = 11, hjust = 0.5)) +
          scale_color_manual(values = c("Pre-Truncation" = "#00CFC1", "Post-Truncation" = "#FC766A"))
        plot_list[[paste(trunc_name, wcol, sep = "_")]] <- p
      } else {
        message(sprintf("Column %s not found in %s", wcol, trunc_name))
      }
    }))
  }
  if (length(plot_list) == 1) {
    return(plot_list[[1]])
  } else {
    return(plot_list)
  }
}
```

**Arguments**

- `trunc_list`: A list of data frames, each containing a dataset with truncated weights.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `weight_cols`: The name of the columns holding the weights.
- `dataset_name`: The dataset name for plot titles.
- `threshold`: A numeric value specifying the acceptable limit for standardized mean differences (default is 0.1).

#### Trimming

`compute_abs_smd_trim()` computes absolute SMDs for multiple trimmed datasets.
```{r}
compute_abs_smd_trim <- function(trimming_list, treat, covar) {
  smd_list <- lapply(names(trimming_list), function(name) {
    data <- trimming_list[[name]]
    bal_obj <- cobalt::bal.tab(
      as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
      data = data,
      un = TRUE, 
      s.d.denom = "treated"
    )
    smd_df <- as.data.frame(bal_obj$Balance)
    smd_vals <- abs(smd_df$Diff.Un)
    mean_smd <- mean(smd_vals, na.rm = TRUE)
    max_smd  <- max(smd_vals, na.rm = TRUE)
    return(data.frame(
      Method   = name,
      Mean_Abs_SMD = mean_smd,
      Max_Abs_SMD  = max_smd 
    ))
  })
  smd_summary <- do.call(rbind, smd_list)
  rownames(smd_summary) <- NULL
  return(smd_summary)
}
```

**Arguments**

- `trimming_list`: A list of data frames, each a trimmed dataset.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.

`compute_ess_trim()` extracts ESS information for multiple trimmed datasets.
```{r}
compute_ess_trim <- function(trimming_list, treat, covar) {
  ess_list <- lapply(trimming_list, function(data) {
    bal_obj <- cobalt::bal.tab(as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
                               data = data, un = TRUE)
    samples <- bal_obj$Observations
    df <- as.data.frame(samples)[c("Control", "Treated")]
    return(df)
  })
  ess_df <- do.call(rbind, ess_list)
  ess_df$Method <- names(trimming_list)
  rownames(ess_df) <- NULL
  ess_df <- ess_df[, c("Method", "Control", "Treated")]
  return(ess_df)
}
```

**Arguments**

- `trimming_list`: A list of data frames, each a trimmed dataset.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.

`plot_trim_overlap()` plots overlap diagnostics for a set of trimmed datasets, visualizing covariate balance before and after matching.
```{r}
plot_trim_overlap <- function(data_list, treat, covar, prefix = NULL, 
                              main.size = 1.1, lab.size = 1, axis.size = 1) {
  par(mfrow = c(2, 3), mar = c(4, 4, 1, 1),
      cex.lab = lab.size, font.lab = 1,
      cex.axis = axis.size, font.axis = 1)
  for (name in names(data_list)) {
    data_obj <- data_list[[name]]
    invisible(capture.output(
      assess_overlap(data_obj, treat = treat, cov = covar)
    ))
    title(main = paste0(prefix, " - ", name), cex.main = main.size, font.main = 1)
  }
}
```

**Arguments**

- `data_list`: A list of data frames to plot.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `prefix`: A string prefix for plot titles.
- `main.size`: Numeric value controlling the font size of main plot titles.
- `lab.size`: Numeric value controlling the font size of axis labels.
- `axis.size`: Numeric value controlling the font size of axis tick marks.

#### Combined methods

`compute_abs_smd_comb()` aggregates absolute SMD statistics across multiple datasets, weighting methods, and trimming techniques.
```{r}
compute_abs_smd_comb <- function(combined_list, treat, covar) {
  smd_list <- lapply(names(combined_list), function(weight_method) {
    method_list <- combined_list[[weight_method]]
    res <- lapply(names(method_list), function(trim_method) {
      data <- method_list[[trim_method]]
      # use equal weights if weight column missing
      if (!"weight" %in% colnames(data)) {
        data$weight <- rep(1, nrow(data))
      }
      bal_obj <- cobalt::bal.tab(
        as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
        data = data,
        weights = data$weight,
        un = TRUE,
        s.d.denom = "treated"
      )
      smd_df <- as.data.frame(bal_obj$Balance)
      smd_vals <- abs(smd_df$Diff.Adj)
      method_name <- paste(weight_method, trim_method, sep = "_")
      data.frame(
        Method = method_name,
        Mean_Abs_SMD = mean(smd_vals, na.rm = TRUE),
        Max_Abs_SMD = max(smd_vals, na.rm = TRUE)
      )
    })
    do.call(rbind, res)
  })
  smd_summary <- do.call(rbind, smd_list)
  rownames(smd_summary) <- NULL
  return(smd_summary)
}
```

**Arguments**

- `combined_list`: A list of data frames grouped by weighting and trimming methods.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.

`compute_ess_comb()` aggregates ESS information across various datasets, weighting methods, and trimming techniques.
```{r}
compute_ess_comb <- function(combined_list, treat, covar) {
  ess_list <- lapply(names(combined_list), function(weight_method) {
    method_list <- combined_list[[weight_method]]
    res <- lapply(names(method_list), function(trim_method) {
      data <- method_list[[trim_method]]
      # use equal weights if weight column missing
      if (!"weight" %in% colnames(data)) {
        data$weight <- rep(1, nrow(data))
      }
      bal_obj <- cobalt::bal.tab(
        as.formula(paste(treat, "~", paste(covar, collapse = " + "))),
        data = data,
        weights = data$weight,
        un = TRUE,
        s.d.denom = "treated"
      )
      samples <- bal_obj$Observations
      if ("Adjusted" %in% rownames(samples)) {
        control_ess <- samples["Adjusted", "Control"]
        treated_ess <- samples["Adjusted", "Treated"]
      } else {
        control_ess <- samples[1, "Control"]
        treated_ess <- samples[1, "Treated"]
      }
      method_name <- paste(weight_method, trim_method, sep = "_")
      data.frame(
        Method = method_name,
        Control = control_ess,
        Treated = treated_ess
      )
    })
    do.call(rbind, res)
  })
  ess_summary <- do.call(rbind, ess_list)
  rownames(ess_summary) <- NULL
  return(ess_summary)
}
```

**Arguments**

- `combined_list`: A list of data frames grouped by weighting and trimming methods.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.

`plot_comb_overlap()` creates plots assessing overlap for combinations of weighting and trimming methods across datasets.
```{r}
plot_comb_overlap <- function(comb_meth_cps = NULL, comb_meth_psid = NULL, treat, covar,
                              prefix_cps = NULL, prefix_psid = NULL,
                              main.size = 1.1, lab.size = 1, axis.size = 1) {
  all_combined_list <- list()
  if (!is.null(comb_meth_cps)) all_combined_list$CPS <- comb_meth_cps
  if (!is.null(comb_meth_psid)) all_combined_list$PSID <- comb_meth_psid
  
  for (ds_name in names(all_combined_list)) {
    combined_list <- all_combined_list[[ds_name]]
    plot_list <- list()
    method_names <- character()
    for (weight_method in names(combined_list)) {
      for (trim_method in names(combined_list[[weight_method]])) {
        full_method_name <- paste(weight_method, trim_method, sep = "_")
        plot_list[[full_method_name]] <- combined_list[[weight_method]][[trim_method]]
        method_names <- c(method_names, full_method_name)
      }
    }
    total_plots <- length(plot_list)
    plots_per_page <- 6
    for (i in seq(1, total_plots, by = plots_per_page)) {
      start_idx <- i
      end_idx <- min(i + plots_per_page - 1, total_plots)
      par(mfrow = c(2, 3), mar = c(4, 4, 1, 1), 
          cex.lab = lab.size, font.lab = 1,
          cex.axis = axis.size, font.axis = 1)
      for (j in start_idx:end_idx) {
        data <- plot_list[[j]]
        prefix <- if (ds_name == "CPS") prefix_cps else prefix_psid
        invisible(capture.output(
          assess_overlap(data, treat = treat, cov = covar) 
        ))
        title(main = paste0(prefix, " - ", method_names[j]), cex.main = main.size, font.main = 1)
      }
      if ((end_idx - start_idx + 1) < plots_per_page) {
        for (k in seq_len(plots_per_page - (end_idx - start_idx + 1))) plot.new()
      }
    }
  }
}
```

**Arguments**

- `comb_meth_cps`: A list combining weighting and trimming results for CPS dataset.
- `comb_meth_psid`: A list combining weighting and trimming results for PSID dataset.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `prefix_cps`: A string prefix for plot titles related to CPS results.
- `prefix_psid`: A string prefix for plot titles related to PSID results.
- `main.size`: Numeric value controlling the font size of main plot titles.
- `lab.size`: Numeric value controlling the font size of axis labels.
- `axis.size`: Numeric value controlling the font size of axis tick marks.

`save_comb_hist()` saves the histograms into a pdf file.
```{r}
save_comb_hist <- function(comb_meth_cps = NULL, comb_meth_psid = NULL, treat, covar,
                           prefix = NULL,
                           prefix_cps = NULL, prefix_psid = NULL,
                           path = "graphs/lalonde") {
  dir.create(path, showWarnings = FALSE, recursive = TRUE)
  print(getwd())
  all_combined_list <- list()
  if (!is.null(comb_meth_cps)) all_combined_list$CPS <- comb_meth_cps
  if (!is.null(comb_meth_psid)) all_combined_list$PSID <- comb_meth_psid
  
  file_index <- 1
  for (ds_name in names(all_combined_list)) {
    combined_list <- all_combined_list[[ds_name]]
    plot_list <- list()
    method_names <- character()
    for (weight_method in names(combined_list)) {
      for (trim_method in names(combined_list[[weight_method]])) {
        full_method_name <- paste(weight_method, trim_method, sep = "_")
        plot_list[[full_method_name]] <- combined_list[[weight_method]][[trim_method]]
        method_names <- c(method_names, full_method_name)
      }
    }
    total_plots <- length(plot_list)
    plots_per_page <- 4
    for (i in seq(1, total_plots, by = plots_per_page)) {
      start_idx <- i
      end_idx <- min(i + plots_per_page - 1, total_plots)
      pdf_file <- file.path(path, sprintf("%s_ov_%d.pdf", prefix, file_index))
      pdf(pdf_file, width = 10, height = 8)
      par(mfrow = c(2, 2), mar = c(4, 4, 2, 1), cex.lab = 1, font.lab = 1, cex.axis = 1, font.axis = 1)
      for (j in start_idx:end_idx) {
        data <- plot_list[[j]]
        prefix_str <- if (ds_name == "CPS") prefix_cps else prefix_psid
        assess_overlap(data, treat = treat, cov = covar)
        title(main = paste0(prefix_str, " - ", method_names[j]), cex.main = 1, font.main = 1)
      }
      if ((end_idx - start_idx + 1) < 4) {
        for (k in seq_len(4 - (end_idx - start_idx + 1))) plot.new()
      }
      dev.off()
      file_index <- file_index + 1
    }
  }
}
```

**Arguments**

- `comb_meth_cps`: A list combining weighting and trimming results for CPS dataset.
- `comb_meth_psid`: A list combining weighting and trimming results for PSID dataset.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `prefix`: A string prefix for plot titles related to the respective dataset or model.
- `prefix_cps`: A string prefix for plot titles related to CPS results.
- `prefix_psid`: A string prefix for plot titles related to PSID results.
- `path`: A string specifying the output directory path where the generated histograms will be saved (default is "graphs/lalonde").

#### Getting top methods and datasets
`combine_results()` merges and summarizes SMD and ESS statistics from various analysis stages of non-plus-datasets into a final comparison table.
```{r}
combine_results <- function(dataset_name) {
  dataset_lower <- tolower(dataset_name)
  # retrieve matching results
  smd_matching <- get(paste0("smd_matchit.", dataset_lower))
  ess_matching <- get(paste0("ess_matchit.", dataset_lower))
  # retrieve trimming results
  smd_trimming <- get(paste0("smd_trim.", dataset_lower))
  ess_trimming <- get(paste0("ess_trim.", dataset_lower))
  # retrieve truncation results
  smd_trunc <- get(paste0("smd_trunc.", dataset_lower))
  ess_trunc <- get(paste0("ess_trunc.", dataset_lower))
  # retrieve weighting results
  smd_weighting <- get(paste0("smd_weight.", dataset_lower))
  ess_weighting <- get(paste0("ess_weight.", dataset_lower))
  # retrieve combined SMD and ESS
  smd_combined_var <- paste0("smd_all_comb_meth.", dataset_lower)
  ess_combined_var <- paste0("ess_all_comb_meth.", dataset_lower)
  smd_combined <- get(smd_combined_var)
  ess_combined <- get(ess_combined_var)
  # combine all SMD results
  smd_all <- do.call(rbind, list(
    smd_matching,
    smd_trimming,
    smd_trunc,
    smd_weighting,
    smd_combined[, c("Method", "Mean_Abs_SMD", "Max_Abs_SMD")]
  ))
  # combine all ESS results
  ess_all <- do.call(rbind, list(
    ess_matching,
    ess_trimming,
    ess_trunc,
    ess_weighting,
    ess_combined[, c("Method", "Control", "Treated")]
  ))
  # merge SMD and ESS results by Method
  final_df <- merge(smd_all, ess_all, by = "Method", all = TRUE)
  # remove dataset suffixes for clean labels
  final_df$Method <- gsub("\\.psid", "", final_df$Method, ignore.case = TRUE)
  final_df$Method <- gsub("\\.cps", "", final_df$Method, ignore.case = TRUE)
  # reset row names
  rownames(final_df) <- NULL
  return(final_df)
}
```

**Arguments**

- `dataset_name`: The dataset name for which to combine SMD and ESS summaries.

`combine_results_plus()` merges and summarizes SMD and ESS statistics from various analysis stages of plus- and non-plus-datasets into a final comparison table.
```{r}
combine_results_plus <- function(dataset_name) {
  dataset_lower <- tolower(dataset_name)
  # retrieve standard results
  smd_matching <- get(paste0("smd_matchit.", dataset_lower))
  ess_matching <- get(paste0("ess_matchit.", dataset_lower))
  smd_trimming <- get(paste0("smd_trim.", dataset_lower))
  ess_trimming <- get(paste0("ess_trim.", dataset_lower))
  smd_trunc <- get(paste0("smd_trunc.", dataset_lower))
  ess_trunc <- get(paste0("ess_trunc.", dataset_lower))
  smd_weighting <- get(paste0("smd_weight.", dataset_lower))
  ess_weighting <- get(paste0("ess_weight.", dataset_lower))
  smd_combined <- get(paste0("smd_all_comb_meth.", dataset_lower))
  ess_combined <- get(paste0("ess_all_comb_meth.", dataset_lower))
  # optionally retrieve plus dataset results if they exist
  smd_combined_plus_name <- paste0("smd_all_comb_meth.", dataset_lower, "_plus")
  ess_combined_plus_name <- paste0("ess_all_comb_meth.", dataset_lower, "_plus")
  smd_combined_plus <- if (exists(smd_combined_plus_name)) get(smd_combined_plus_name) else NULL
  ess_combined_plus <- if (exists(ess_combined_plus_name)) get(ess_combined_plus_name) else NULL
  # combine all SMD results
  smd_list <- list(
    smd_matching,
    smd_trimming,
    smd_trunc,
    smd_weighting,
    smd_combined[, c("Method", "Mean_Abs_SMD", "Max_Abs_SMD")]
  )
  if (!is.null(smd_combined_plus)) {
    smd_list <- append(smd_list, list(smd_combined_plus[, c("Method", "Mean_Abs_SMD", "Max_Abs_SMD")]))
  }
  smd_all <- do.call(rbind, smd_list)
  # combine all ESS results
  ess_list <- list(
    ess_matching,
    ess_trimming,
    ess_trunc,
    ess_weighting,
    ess_combined[, c("Method", "Control", "Treated")]
  )
  if (!is.null(ess_combined_plus)) {
    ess_list <- append(ess_list, list(ess_combined_plus[, c("Method", "Control", "Treated")]))
  }
  ess_all <- do.call(rbind, ess_list)
  # merge SMD and ESS results
  final_df <- merge(smd_all, ess_all, by = "Method", all = TRUE)
  # clean method names
  final_df$Method <- gsub("\\.psid|\\.cps", "", final_df$Method, ignore.case = TRUE)
  rownames(final_df) <- NULL
  return(final_df)
}
```

**Arguments**

- `dataset_name`: The dataset name for which to combine SMD and ESS summaries.

`save_csv()` saves the tables into a CSV file.
```{r}
save_csv <- function(data, filename) {
  folder <- "tables"
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  file_csv <- file.path(folder, paste0(filename, ".csv"))
  write.csv(data, file = file_csv, row.names = FALSE)
}
```

**Arguments**

- `data`: The dataset.
- `filename`: The name of the file that is to be saved as CSV.

`assess_methods()` scores and ranks methods based on a weighted combination of balance and sample size metrics.
```{r}
assess_methods <- function(data) {
  data %>%
    mutate(
      # ess score
      ess_balance_ratio = pmin(Control, Treated) / pmax(Control, Treated),
      ess_total = Control + Treated,
      # normalize balance_ratio and ess_total
      ess_balance_score = (ess_balance_ratio - min(ess_balance_ratio, na.rm = TRUE)) /
        (max(ess_balance_ratio, na.rm = TRUE) - min(ess_balance_ratio, na.rm = TRUE)),
      ess_size_score = (ess_total - min(ess_total, na.rm = TRUE)) /
        (max(ess_total, na.rm = TRUE) - min(ess_total, na.rm = TRUE)),
      # combine size and balance equally
      ess_score = 0.5 * ess_balance_score + 0.5 * ess_size_score,
      # smd score
      smd_score = 1 - (Mean_Abs_SMD - min(Mean_Abs_SMD, na.rm = TRUE)) /
        (max(Mean_Abs_SMD, na.rm = TRUE) - min(Mean_Abs_SMD, na.rm = TRUE)),
      # composite score
      Score = 0.5 * smd_score + 0.5 * ess_score
    ) %>%
    dplyr::select(Method, Score) %>%
    arrange(desc(Score))
}
```

**Argument**

- `data`: A data frame containing method performance metrics.

`get_top_methods()` retrieves the top-performing methods based on combined scoring criteria.
```{r}
get_top_methods <- function(summary_df, top_n = 5) {
  if (!all(c("Method", "Score") %in% names(summary_df))) {
    stop("Data frame must contain columns 'Method' and 'Score'")
  }
  top_methods_df <- summary_df %>%
    arrange(desc(Score)) %>%
    head(top_n)
  return(top_methods_df$Method)
}
```

**Arguments**

- `summary_df`: A data frame including columns used for ranking.
- `top_n`: Integer specifying the number of top-performing methods to return (default 5).

`create_top5_datasets()` extracts datasets corresponding to the top five ranked methods for further analysis or presentation.
```{r}
create_top5_datasets <- function(dataset_list, top5_method_names) {
  lapply(top5_method_names, function(method_name) {
    if (!method_name %in% names(dataset_list)) {
      stop(paste0("Method '", method_name, "' not found in the dataset lookup list"))
    }
    ds <- dataset_list[[method_name]]
    if (inherits(ds, "matchit")) {
      # for matchit class, extract matched data
      return(as.data.frame(match.data(ds)))
    } else if (is.data.frame(ds)) {
      # if already dataframe, return as is
      return(ds)
    } else if (is.vector(ds)) {
      # merge with original data
      if (!"original" %in% names(dataset_list)) {
        stop("Original dataset not found in dataset_list to merge weights")
      }
      original_df <- dataset_list[["original"]]
      if (length(ds) != nrow(original_df)) {
        stop(paste0("Length of weight vector for method '", method_name, "' does not match number of rows in original dataset"))
      }
      # create new data frame with weights appended
      weighted_df <- original_df
      weighted_df[["weights"]] <- ds
      return(weighted_df)
    } else {
      stop(paste("Unsupported data type for method", method_name))
    }
  })
}
```

**Arguments**

- `dataset_list`: A list of datasets or MatchIt objects containing processed results for different methods.
- `top5_method_names`: A character vector of method names corresponding to the top-ranked datasets.

`save_top5_individual_files()` saves datasets of the top five methods as individual files for reproducibility or sharing. 
```{r}
save_top5_datasets <- function(combined_methods_list, top5_method_names, prefix) {
  for (i in seq_along(top5_method_names)) {
    method_name <- top5_method_names[i]
    if (!method_name %in% names(combined_methods_list)) {
      warning(paste0("Method '", method_name, "' not found in combined methods list"))
      next
    }
    dataset_to_save <- combined_methods_list[[method_name]]
    file_name <- sprintf("tutorial/data/top%d_%s_method_%s.RData", i, prefix, method_name)
    save(dataset_to_save, file = file_name)
  }
}
```

**Arguments**

- `combined_methods_list`: A list of all combined method results.
- `top5_method_names`: A character vector of method names corresponding to the top-ranked datasets.
- `prefix`: A string prefix used for saved file names to identify dataset.

<br>

### Estimating 
#### Average treatment effect of the treated (ATT)

<a name="est-and-plot-att-catt-qtet"></a>

`estimate_all()` runs multiple treatment effect estimators on a dataset and returns point estimates, standard errors and confidence intervals.***
```{r}
# difference in means
# diff()
diff <- function(data, Y, treat) {
  fml <- as.formula(paste(Y, "~", treat))
  out <- summary(lm_robust(fml, data = data, se_type = "stata"))$coefficients[treat, c(1, 2, 5, 6)]
  return(out) # extract coef, se, ci.lower, ci.upper
}

# regression adjustment
# reg()
reg <- function(data, Y, treat, covar) {
  fml <- as.formula(paste(Y, "~", treat, "+", paste(covar, collapse = " + ")))
  out <- summary(lm_robust(fml, data = data, se_type = "stata"))$coefficients[treat, c(1, 2, 5, 6)]
  # extract coef, se, ci.lower, ci.upper
  return(out)
}

# matching
# library(Matching)
# matching()
matching <- function(data, Y, treat, covar) {
  m.out <- Match(Y = data[, Y], Tr = data[, treat], X = data[, covar], Z = data[, covar],
                 estimand = "ATT", M = 5, replace = TRUE, ties = TRUE, BiasAdjust = TRUE)
  out <- c(m.out$est[1], m.out$se[1], m.out$est[1] - 1.96 * m.out$se[1],
           m.out$est[1] + 1.96 * m.out$se[1])
  return(out)
}

# psm()
psm <- function(data, Y, treat, covar) {
  ps <- probability_forest(X = data[, covar],
                           Y = as.factor(data[,treat]), seed = 42, num.trees = 4000)$predictions[,2]
  m.out <- Match(Y = data[, Y], Tr = data[, treat], X = matrix(ps, nrow(data), 1),
                 estimand = "ATT", M = 5, replace = TRUE, ties = FALSE, BiasAdjust = FALSE)
  if (is.null(m.out$se)==FALSE) {
    se <- m.out$se[1]
  } else {
    se <- m.out$se.standard[1]
  }
  out <- c(m.out$est[1], se, m.out$est[1] - 1.96 * se,
           m.out$est[1] + 1.96 * se)
  return(out)
}

# OM (reg)
# om.reg()
om.reg <- function(data, Y, treat, covar) {
  tr <- which(data[, treat] == 1)
  co <- which(data[, treat] == 0)
  fml <- as.formula(paste(Y, "~", paste(covar, collapse = " + ")))
  out.co <- lm(fml, data = data[co, ])
  Y.tr.hat <- predict(out.co, newdata = data[tr, covar, drop = FALSE])
  newdata <- cbind.data.frame(Y = c(data[tr, Y], Y.tr.hat), treat = rep(c(1, 0), each = length(tr)))
  out <- summary(lm_robust(Y ~ treat, data = newdata, se_type = "stata"))$coefficients["treat", c(1, 2, 5, 6)]
  return(out)
}

# OM (grf)
# library(grf)
# om.grf()
om.grf <- function(data, Y, treat, covar) {
  tr <- which(data[, treat] == 1)
  co <- which(data[, treat] == 0)
  out.co <- regression_forest(X = data[co, covar, drop = FALSE], Y = as.vector(data[co, Y]) )
  Y.tr.hat <- as.vector(unlist(predict(out.co, newdata = data[tr, covar, drop = FALSE])))
  newdata <- cbind.data.frame(Y = c(data[tr, Y], Y.tr.hat), treat = rep(c(1, 0), each = length(tr)))
  out <- summary(lm_robust(Y ~ treat, data = newdata, se_type = "stata"))$coefficients["treat", c(1, 2, 5, 6)]
  return(out)
}

# IPW
# ipw()
ipw <- function(data, Y, treat, covar) {
  ps <- probability_forest(X = data[, covar, drop = FALSE], Y = as.factor(data[, treat]), seed = 42)$predictions[,2]
  fml <- as.formula(paste(Y, "~", treat))
  weights <- rep(1, nrow(data))
  co <- which(data[, treat] == 0)
  weights[co] <- ps[co]/(1-ps[co])
  out <- summary(lm_robust(fml, data = data, weights = weights, se_type = "stata"))$coefficients[treat, c(1, 2, 5, 6)]
  # extract coef, se, ci.lower, ci.upper
  return(out)
}

# CBPS
# library("CBPS")
# cbps()
cbps <- function(data, Y, treat, covar) {
  fml <- as.formula(paste(treat, "~", paste(covar, collapse = " + ")))
  ps <- quiet(CBPS(fml, data = data, standardize = TRUE)$fitted.values)
  fml <- as.formula(paste(Y, "~", treat))
  weights <- rep(1, nrow(data))
  co <- which(data[, treat] == 0)
  weights[co] <- ps[co]/(1-ps[co])
  out <- summary(lm_robust(fml, data = data, weights = weights, se_type = "stata"))$coefficients[treat, c(1, 2, 5, 6)]
  return(out)
}

# ebal()
#library(hbal)
ebal <- function(data, Y, treat, covar) {
  ebal.out <- hbal::hbal(Y = Y, Treat = treat, X = covar,  data = data, expand.degree = 1)
  out <- hbal::att(ebal.out, dr = FALSE)[1, c(1, 2, 5, 6)]
  return(out)
}

# hbal
# hbal <- function(data, Y, treat, covar) {
#   hbal.out <- hbal::hbal(Y = Y, Treat = treat, X = covar,  data = data, expand.degree = 2, # cv = TRUE)
#   out <- hbal::att(hbal.out, dr = FALSE)[1, c(1, 2, 5, 6)]
#   return(out)
# }

# aipw_grf()
aipw <- function(data, Y, treat, covar) {
  tryCatch({
    #library("grf")
    for (var in c(Y, treat, covar)) {
      data[, var] <- as.vector(data[, var])
    }
    c.forest <- causal_forest(X = data[, covar, drop = FALSE], Y = data[, Y],
                              W = data[, treat], seed = 42)
    att <- average_treatment_effect(c.forest, target.sample = "treated", method = "AIPW")
    att <- c(att, att[1] - 1.96 * att[2], att[1] + 1.96 * att[2])
    return(att)
  }, error = function(e) {
    cat("Error in aipw_grf method:", e$message, "\n")
    return(c(NA, NA, NA, NA))
  })
}

# aipw.match()
aipw.match <- function(data, Y, treat, covar) { # match on ps
  ps <- probability_forest(X = data[, covar], Y = as.factor(data[, treat]), seed = 42)$predictions[,2]
  m.out <- Match(Y = data[, Y], Tr = data[, treat], X = ps,
                 estimand = "ATT", M = 1, replace = FALSE, ties = FALSE, BiasAdjust = FALSE)
  mb <- quiet(MatchBalance(treat ~ ps, data = data, match.out = m.out, nboots= 0))
  ks <- mb$AfterMatching[[1]]$ks$ks$statistic
  s <- data[c(m.out$index.treated, m.out$index.control), ]
  out <- aipw(s, Y, treat, covar)
  # return(out)
  return(c(out, ks))
}

# aipw_ow() ***
aipw_ow <- function(data, Y, treat, covar) {
  tryCatch({
    for (var in c(Y, treat, covar)) {
      data[, var] <- as.vector(data[, var])
    }
    X <- data[, covar, drop = FALSE]
    Y <- data[, Y]
    W <- data[, treat]
    # run dml_with_smoother with AIPW_ATT
    dml_fit <- dml_with_smoother(Y = Y, D = W, X = X,
                                 estimators = c("AIPW_ATT"),
                                 smoother = "honest_forest",
                                 n_cf_folds = 5,
                                 n_reps = 1)
    # extract estimate and SE from summary
    summ <- summary(dml_fit, quiet = TRUE)
    est <- summ["AIPW-ATT", "Estimate"]
    se <- summ["AIPW-ATT", "SE"]
    ci_lower <- est - 1.96 * se
    ci_upper <- est + 1.96 * se
    return(c(est, se, ci_lower, ci_upper))
  }, error = function(e) {
    cat("Error in aipw_ow method:", e$message, "\n")
    return(c(NA, NA, NA, NA))
  })
}

### This script checks for robustness by estimating original model
### using double/debiased machine learning using DoubleML package
dml <-function(data, Y = NULL, treat = NULL, covar = NULL, clust_var = NULL, ml_l = lrn("regr.lm"), ml_m = lrn("regr.lm")){
  tryCatch({
    if(is.null(covar)){
      stop("No controls in specification.")
    }
    #require(DoubleML)
    #require(mlr3learners)
    #require(fixest)
    #require(ggplot2)
    if(is.null(clust_var) == TRUE){
      dat = data[,c(Y,treat,covar)]
      dat = na.omit(dat)
      dml_dat = DoubleMLData$new(dat,
                                 y_col = Y,
                                 d_cols = treat,
                                 use_other_treat_as_covariate = FALSE,
                                 x_cols = covar)
    }else{
      dat = data[,c(Y, treat, covar, clust_var)]
      dat[,clust_var] = as.numeric(factor(dat[,clust_var]))
      dat = dat[is.na(dat[,Y]) == FALSE,]
      dat = dat[is.na(dat[,D]) == FALSE,]
      features = data.frame(model.matrix(formula(paste(c('~ 1',treat,covar), collapse="+")), dat))
      dat = cbind(dat[,c(Y,clust_var)],features)
      dml_dat = DoubleMLClusterData$new(dat,
                                        y_col = Y,
                                        d_cols = treat,
                                        cluster_cols = clust_var,
                                        use_other_treat_as_covariate = FALSE,
                                        x_cols = covar)
    }
    # set active treatment treatment
    dml_dat$set_data_model(treat)
    # estimate with DML
    set.seed(pi)
    dml_mod = DoubleMLPLR$new(dml_dat, ml_l=ml_l, ml_m=ml_m)
    quiet(dml_mod$fit())
    out = c(dml_mod$coef[treat], dml_mod$se[treat], dml_mod$confint()[treat,])
    return(out)
  }, error = function(e) {
    cat("Error in dml method:", e$message, "\n")
    return(c(NA, NA, NA, NA))
  })
}

# estimate_all ***
estimate_all <- function(data, Y, treat, covar, 
                         methods = c("diff", "reg", "om.reg", "om.grf",
                                     "matching", "psm", "ipw", "cbps", "ebal", 
                                     "dml", "aipw_grf", "aipw_ow")) {
  
  results <- as.data.frame(matrix(NA, length(methods), 4))
  rownames(results) <- methods
  colnames(results) <- c("Estimate", "SE", "CI_lower", "CI_upper")
  m <- 1
  if ("diff" %in% methods) {
    results[m, ] <- diff(data, Y, treat) 
    m <- m + 1
  }
  if ("reg" %in% methods) {
    results[m, ] <- reg(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("om.reg" %in% methods) {
    results[m, ] <- om.reg(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("om.grf" %in% methods) {
    results[m, ] <- om.grf(data, Y, treat, covar) 
    m <- m + 1
  } 
  if ("matching" %in% methods) {
    results[m, ] <- matching(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("psm" %in% methods) {
    results[m, ] <- psm(data, Y, treat, covar) 
    m <- m + 1
  }  
  if ("ipw" %in% methods) {
    results[m, ] <- ipw(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("cbps" %in% methods) {
    results[m, ] <- cbps(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("ebal" %in% methods) {
    results[m, ] <- quiet(ebal(data, Y, treat, covar))
    m <- m + 1
  }
  # if ("hbal" %in% methods) {
  #   results[m, ] <- quiet(hbal(data, Y, treat, covar))
  #   m <- m + 1
  # }
  if ("dml" %in% methods) {
    results[m, ] <-dml(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("aipw_grf" %in% methods) {
    results[m, ] <- aipw(data, Y, treat, covar) 
    m <- m + 1
  }
  if ("aipw_ow" %in% methods) {
    results[m, ] <- aipw_ow(data, Y, treat, covar)
    m <- m + 1
  }
  return(results)
}
```

**Function calls**

-   `quiet`: Suppresses output from a function call.
-   `diff`: Difference in means estimator. It runs a linear regression adjusting for robust standard errors and returns the coefficient, standard error, and confidence interval for the treatment variable.
-   `reg`: Regression adjustment. Similar to `diff` but includes additional covariates in the regression model.
-   `matching`: Propensity score matching using the Matching package. It aligns treated units to control units based on covariates and returns the estimated ATT and its confidence interval.
-   `psm`: Propensity score matching using a probability forest, followed by matching and estimation of the ATT.
-   `om.reg`: Outcome modeling using regression. It predicts the outcome for the treated units based on the model fitted to the control units, and then estimates the ATT.
-   `om.grf`: Outcome modeling using generalized random forests, noted as GRF.
-   `ipw`: Inverse probability weighting,denoted as IPW. It weights observations by the inverse of their estimated propensity scores and calculates the treatment effect with a weighted regression.
-   `cbps`: Covariate balancing propensity score, noted as CBPS. It estimates propensity scores to achieve balance on covariates across groups.
-   `ebal`: Entropy balancing. It reweights the data to balance the covariate distributions.
-   `hbal`: Hierarchical balancing. It is an extension of `ebal` with more complex balancing methods.
-   `aipw`: Augmented inverse probability weighting, noted as AIPW.
-   `aipw.match`: Combines matching on propensity scores with AIPW.
-   `aipw_ow`: AIPW computed via outcome weights.
-   `dml`: Double machine learning. It uses machine learning algorithms to control for confounders when estimating treatment effects.

`plot_coef()` plots treatment effect point estimates and corresponding confidence intervals comparatively across various estimators.***
```{r}
plot_coef <- function(out, 
                      methods = c("diff", "reg", "om.reg", "om.grf", 
                                  "matching", "psm", "ipw", "cbps", "ebal", 
                                  "dml", "aipw_grf", "aipw_ow"),
                      labels = c("Diff-in-Means", "Reg", "OM: Reg", "OM: GRF",
                                 "NN\nMatching", "PS\nMatching",
                                 "IPW", "CBPS", "Ebal", "DML\nElasnet", "AIPW-GRF", "AIPW-OW"),
                      main = NULL,
                      ylab = "Estimate",
                      band = NULL,
                      line = NULL,
                      grid = TRUE,
                      main.pos = 1,
                      main.line = -2,
                      ylim = NULL,
                      textsize = 0.8
) {
  if (is.null(methods) == TRUE) {
    methods <- rownames(out)
  }
  if (is.null(labels) == TRUE) {
    labels <- methods
  }
  # # check
  # if (is.null(out)==FALSE) {
  #   if (inherits(out, "ivDiag") == FALSE) {stop("\"out\" needs to be a \"ltz\" object.")}
  # }
  # 
  # # title
  # if (is.null(main)==TRUE) {
  #   main <- "Estimates with 95% CIs"
  # }
  # data for the plot
  data <- out
  rg <- range(data[,c(3,4)], na.rm = TRUE)
  adj <- rg[2] - rg[1]
  if (is.null(ylim) == TRUE) {
    ylim  <- c(min(0, rg[1] - 0.3*adj), max(0, rg[2] + 0.35*adj))
  }
  adj2 <- ylim[2] - ylim[1] 
  # Set up the plot
  ncoefs <- length(methods)
  par(mar = c(2.5, 4, 1, 2))
  plot(1: ncoefs, data[, 1], xlim = c(0.5, ncoefs + 0.5), ylim = ylim,
       ylab = "", xlab = "", main = "", 
       axes = FALSE, xaxt = "n", yaxt = "n", type = "n")
  axis(1, at = 1: ncoefs, labels =  labels, las = 1, cex.axis = 0.8)
  axis(2, cex.axis = 0.7)
  mtext(main, main.pos, line = main.line, cex = textsize)
  mtext(ylab, 2, line = 2.5)
  if (is.null(band) == FALSE) {
    rect(-0.5, band[1], ncoefs + 1, band[2], col = "#ff000030", border = "white") # label at bottom
  }
  if (is.null(line) == FALSE) {
    abline(h = line, col = "red", lty = 2)
  }
  if (grid == TRUE) {
    abline(h = axTicks(2), lty = "dotted", col = "gray50")
    abline(v = c(0.5, c(1: ncoefs) + 0.5), lty = "dotted", col = "gray50") # horizontal grid
  }
  abline(h = 0, col = "red", lwd = 2, lty = "solid")
  segments(y0 = data[, 3], x0 = c(1: ncoefs), y1 = data[, 4], x1 = c(1: ncoefs), lwd = 2) #CI
  points(1: ncoefs, data[, 1], pch = 16, col = 1, cex = 1.2) #point coefs
  box()
}
```

`save_att_panels()` generates and saves paginated pdf plots that display ATT estimates and their confidence bands from multiple methods.
```{r}
save_att_panels <- function(out_list, plot_titles, band_list, est_list,
                              prefix = "ldw_model_a_plus",
                              plots_per_page = 4,
                              ylim = c(-15500, 5500),
                              folder = "graphs/lalonde") {
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  n <- length(out_list)
  pages <- ceiling(n / plots_per_page)
  
  for (p in seq_len(pages)) {
    start <- (p - 1) * plots_per_page + 1
    end <- min(p * plots_per_page, n)
    file_name <- file.path(folder, paste0(prefix, "_", p, ".pdf"))
    pdf(file_name, width = 8, height = 11)
    par(mfrow = c(plots_per_page, 1), mar = c(3, 4, 3, 2))
    for (i in start:end) {
      plot_coef(out_list[[i]], band = band_list[[i]], line = est_list[[i]],
                ylim = ylim, main = plot_titles[i])
    }
    dev.off()
  }
}
```

**Arguments**

- `out_list`: A list of output objects containing ATT estimates and confidence intervals for different methods.
- `plot_titles`: A character vector of titles for each plot.
- `band_list`: A list of numeric vectors specifying the confidence interval bounds for each plot.
- `est_list`: A list or numeric vector representing the point estimates (ATT values) for each plot.
- `prefix`: A string prefix used for saved file names to identify dataset.
- `plots_per_page`: The number of plots to display per page / panel (default 4).
- `ylim`: A numeric vector specifying the y-axis limits for the plots.
- `folder`: A string specifying the folder path where the PDF panels will be saved (default is "graphs/lalonde").

`create_matrix_results()` organizes estimation results into a formatted matrix suitable for reporting or tables.
```{r}
create_matrix_results <- function(all_outs, sample_names) {
  n_samples <- length(sample_names)
  n_estimators <- nrow(all_outs[[1]])
  # initialize empty matrix
  result_mat <- matrix("", nrow = n_estimators + 1, ncol = n_samples * 2)
  # alternating column names (estimate / SE)
  cnames <- character(n_samples * 2)
  for (j in seq_along(sample_names)) {
    cnames[(j - 1) * 2 + 1] <- sample_names[j]
    cnames[(j - 1) * 2 + 2] <- ""
  }
  colnames(result_mat) <- cnames
  estimator_names <- rownames(all_outs[[1]])
  rownames(result_mat) <- c("Experimental Benchmark", estimator_names)
  # extract benchmark estimates 
  bench.exp  <- all_outs[[1]][1, 1:2]   # experimental LDW
  bench.cps  <- all_outs[[2]][1, 1:2]   # LDW-CPS1 trimmed benchmark
  bench.psid <- all_outs[[3]][1, 1:2]   # LDW-PSID1 trimmed benchmark
  # fill results column by column
  for (j in seq_along(all_outs)) {
    out <- all_outs[[j]]
    # assign benchmark row depending on dataset type
    if (j <= 3) {  
      result_mat[1, (j - 1) * 2 + 1] <- sprintf("%.2f", bench.exp[1])
      result_mat[1, (j - 1) * 2 + 2] <- paste0("(", sprintf("%.2f", bench.exp[2]), ")")
    } 
    else if (grepl("CPS", sample_names[j], ignore.case = TRUE)) {
      result_mat[1, (j - 1) * 2 + 1] <- sprintf("%.2f", bench.cps[1])
      result_mat[1, (j - 1) * 2 + 2] <- paste0("(", sprintf("%.2f", bench.cps[2]), ")")
    } 
    else if (grepl("PSID", sample_names[j], ignore.case = TRUE)) {
      result_mat[1, (j - 1) * 2 + 1] <- sprintf("%.2f", bench.psid[1])
      result_mat[1, (j - 1) * 2 + 2] <- paste0("(", sprintf("%.2f", bench.psid[2]), ")")
    }
    # fill in estimates + SEs
    for (i in 2:(n_estimators + 1)) {
      result_mat[i, (j - 1) * 2 + 1] <- sprintf("%.2f", out[i - 1, 1])
      result_mat[i, (j - 1) * 2 + 2] <- paste0("(", sprintf("%.2f", out[i - 1, 2]), ")")
    }
  }
  return(result_mat)
}
```

**Arguments**

- `all_outs`: A list of output objects containing ATT estimates and confidence intervals for different methods.
- `sample_names`: A character vector of sample / dataset names corresponding to `all_outs`.

`eval_att()` computes summary statistics of the treatment effect estimates. 
```{r}
eval_att <- function(result) {
  data.frame(
    Mean_SE = mean(result[, "SE"], na.rm = TRUE), # mean standard error 
    Min_Estimate = min(result[, "Estimate"], na.rm = TRUE), # maximum estimate 
    Max_Estimate = max(result[, "Estimate"], na.rm = TRUE), # minimum estimate 
    Diff_Estimate = max(result[, "Estimate"], na.rm = TRUE) - min(result[, "Estimate"], na.rm = TRUE)
  )
}
```

**Argument**

- `result`: A data frame containing treatment effect estimation results.

#### Conditional average treatment effect on the treated (CATT)

`plot_catt_panels()` visualizes CATT comparing multiple methods.
```{r}
plot_catt_panels <- function(exp_catt, catt_list, plot_titles, plots_per_page = 4, range = c(-8000, 8000)) {
  n <- length(catt_list)
  num_pages <- ceiling(n / plots_per_page)
  catt_ldw <- exp_catt$catt
  att_ldw  <- exp_catt$att[1]
  id_ldw   <- if (!is.null(exp_catt$id)) exp_catt$id else seq_along(catt_ldw)
  for (page in seq_len(num_pages)) {
    start_idx <- (page - 1) * plots_per_page + 1
    end_idx   <- min(page * plots_per_page, n)
    par(mfrow = c(2, 2), mar = c(4.5, 5, 3, 2))
    for (i in start_idx:end_idx) {
      catt2 <- catt_list[[i]]$catt
      att2  <- catt_list[[i]]$att[1]
      id2   <- if (!is.null(catt_list[[i]]$id)) catt_list[[i]]$id else seq_along(catt2)
      common_ids <- intersect(id_ldw, id2)
      idx_ldw    <- match(common_ids, id_ldw)
      idx_other  <- match(common_ids, id2)
      catt1_plot <- catt_ldw[idx_ldw]
      catt2_plot <- catt2[idx_other]
      plot_catt(
        catt1 = catt1_plot,
        catt2 = catt2_plot,
        att1  = att_ldw,
        att2  = att2,
        xlab  = "CATT (Experimental)",
        ylab  = plot_titles[i],
        main  = plot_titles[i],
        axes.range = range
      )
    }
  }
}
```

**Arguments**

- `exp_catt`: An object containing CATT estimates used as experimental benchmark.
- `catt_list`: A list of objects containing CATT estimates from each comparison with the experimental benchmark.
- `plot_titles`: A character vector of titles for each plot.
- `plots_per_page`: The number of plots to display per page / panel (default 4).
- `range`: A numeric vector of axis limits for plots (default c(-8000, 8000)).

`save_catt_panels()`generates and saves multi-page PDF plots comparing CATT estimates across different methods.
```{r}
save_catt_panels <- function(
    exp_catt, 
    catt_list, 
    plot_titles, 
    prefix = "catt_top5", 
    plots_per_page = 4, 
    range = c(-8000, 8000), 
    folder = "graphs/lalonde"
) {
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  n <- length(catt_list)
  num_pages <- ceiling(n / plots_per_page)
  catt_ldw <- exp_catt$catt
  att_ldw  <- exp_catt$att[1]
  id_ldw   <- if (!is.null(exp_catt$id)) exp_catt$id else seq_along(catt_ldw)
  for (page in seq_len(num_pages)) {
    start_idx <- (page - 1) * plots_per_page + 1
    end_idx   <- min(page * plots_per_page, n)
    file_name <- file.path(folder, paste0(prefix, "_", page, ".pdf"))
    pdf(file = file_name, width = 10, height = 12)
    par(mfrow = c(plots_per_page, 1), mar = c(4.5, 5, 3, 2))
    for (i in start_idx:end_idx) {
      catt2 <- catt_list[[i]]$catt
      att2  <- catt_list[[i]]$att[1]
      id2   <- if (!is.null(catt_list[[i]]$id)) catt_list[[i]]$id else seq_along(catt2)
      common_ids <- intersect(id_ldw, id2)
      idx_ldw    <- match(common_ids, id_ldw)
      idx_other  <- match(common_ids, id2)
      catt1_plot <- catt_ldw[idx_ldw]
      catt2_plot <- catt2[idx_other]
      plot_catt(
        catt1 = catt1_plot,
        catt2 = catt2_plot,
        att1  = att_ldw,
        att2  = att2,
        xlab  = "CATT (Experimental)",
        ylab  = plot_titles[i],
        main  = plot_titles[i],
        axes.range = range
      )
    }
    # fill empty panels if last page is not full
    if ((end_idx - start_idx + 1) < plots_per_page) {
      for (k in seq_len(plots_per_page - (end_idx - start_idx + 1))) plot.new()
    }
    dev.off()
  }
}
```

**Arguments**

- `exp_catt`: An object containing CATT estimates used as experimental benchmark.
- `catt_list`: A list of objects containing CATT estimates from each comparison with the experimental benchmark.
- `plot_titles`: A character vector of titles for each plot.
- `prefix`: A string prefix for saved PDF filenames.
- `plots_per_page`: The number of plots to display per page / panel (default 4).
- `range`: A numeric vector of axis limits for plots (default c(-8000, 8000)).
- `folder`: A string specifying the folder path where the PDF panels will be saved (default is "graphs/lalonde").

`save_main_catt_panels()`
```{r}
save_main_catt_panels <- function(
    catt_refs, catt_comps, ylabels,
    prefix = "catt_main",
    plots_per_page = 4,
    main_titles = NULL,
    range = c(-8000, 8000),
    folder = "graphs/lalonde"
) {
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  n_panels <- length(catt_comps)
  num_pages <- ceiling(n_panels / plots_per_page)
  for (page in seq_len(num_pages)) {
    start_idx <- (page - 1) * plots_per_page + 1
    end_idx <- min(page * plots_per_page, n_panels)
    file_name <- file.path(folder, paste0(prefix, "_", page, ".pdf"))
    pdf(file = file_name, width = 10, height = 3 * plots_per_page)
    par(mfrow = c(plots_per_page, 1), mar = c(4.5, 5, 3, 2))
    for (i in start_idx:end_idx) {
      ref_idx <- min(i, length(catt_refs))
      catt1 <- catt_refs[[ref_idx]]$catt
      catt2 <- catt_comps[[i]]$catt
      att1  <- catt_refs[[ref_idx]]$att[1]
      att2  <- catt_comps[[i]]$att[1]
      plot_catt(
        catt1 = catt1,
        catt2 = catt2,
        att1  = att1,
        att2  = att2,
        xlab  = "CATT (Reference)",
        ylab  = ylabels[i],
        main  = if (is.null(main_titles)) paste("Comparison", i) else main_titles[i],
        axes.range = range
      )
    }
    # fill empty panels if last page not full
    if ((end_idx - start_idx + 1) < plots_per_page) {
      for (k in seq_len(plots_per_page - (end_idx - start_idx + 1))) plot.new()
    }
    dev.off()
  }
}
```

**Arguments**

- `catt_refs`: List of reference CATT objects used for x-axis values in each comparison plot.
- `catt_comps`: List of comparison CATT objects used for the y-axis in each plot.
- `ylabels`: Character vector of plot y-axis labels.
- `prefix`: A string prefix for saved PDF filenames.
- `plots_per_page`: The number of plots to display per page / panel (default 4).
- `main_titles`: Character vector of main titles to appear at the top of each plot.
- `range`: A numeric vector of axis limits for plots (default c(-8000, 8000)).
- `folder`: 

`save_plus_catt_panels()`
```{r}
save_plus_catt_panels <- function(
    catt1_list, catt2_list,
    ylabels,
    prefix = "catt_plus",
    plots_per_page = 4,
    main_titles = NULL,
    range = c(-8000, 8000),
    folder = "graphs/lalonde"
) {
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  n_panels <- min(length(catt1_list), length(catt2_list))
  num_pages <- ceiling(n_panels / plots_per_page)
  for (page in seq_len(num_pages)) {
    start_idx <- (page - 1) * plots_per_page + 1
    end_idx <- min(page * plots_per_page, n_panels)
    file_name <- file.path(folder, paste0(prefix, "_", page, ".pdf"))
    pdf(file = file_name, width = 10, height = 3 * plots_per_page)
    par(mfrow = c(plots_per_page, 1), mar = c(4.5, 5, 3, 2))
    for (i in start_idx:end_idx) {
      catt1 <- catt1_list[[i]]$catt
      catt2 <- catt2_list[[i]]$catt
      att1  <- catt1_list[[i]]$att[1]
      att2  <- catt2_list[[i]]$att[1]
      plot_catt(
        catt1 = catt1,
        catt2 = catt2,
        att1  = att1,
        att2  = att2,
        xlab  = "CATT (Reference/Trimmed)",
        ylab  = ylabels[i],
        main  = if (is.null(main_titles)) paste("Trimmed Panel", i) else main_titles[i],
        axes.range = range
      )
    }
    # fill empty panels if last page not full
    if ((end_idx - start_idx + 1) < plots_per_page) {
      for (k in seq_len(plots_per_page - (end_idx - start_idx + 1))) plot.new()
    }
    dev.off()
  }
}
```

**Arguments**

- `catt1_list`: List of CATT objects for x-axis values.
- `catt2_list`: List of CATT objects for y-axis values.
- `ylabels`: Character vector of plot y-axis labels.
- `prefix`: A string prefix for saved PDF filenames.
- `plots_per_page`: The number of plots to display per page / panel (default 4).
- `main_titles`: Character vector of main titles to appear at the top of each plot.
- `range`: A numeric vector of axis limits for plots (default c(-8000, 8000)).
- `folder`:

`eval_catt()` computes summary statistics of CATT vectors for each method, returning the minimum, maximum, mean, and range (difference) of the CATT estimates.
```{r}
eval_catt <- function(all_catt, plot_titles) {
  do.call(rbind, lapply(seq_along(all_catt), function(i) {
    catt_vec <- all_catt[[i]]$catt
    data.frame(
      Method = plot_titles[i],
      Min_Catt = min(catt_vec, na.rm = TRUE),
      Max_Catt = max(catt_vec, na.rm = TRUE),
      Mean_Catt = mean(catt_vec, na.rm = TRUE),
      Diff_Catt = max(catt_vec, na.rm = TRUE) - min(catt_vec, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  }))
}
```

**Arguments**

- `all_catt`: A list of objects containing conditional average treatment effect on treated (CATT) estimates.
- `plot_titles`: A character vector of titles for each plot, corresponding to the outputs in `all_catt`.

#### Quantile treatment effect on treated (QTET)

`plot_qte_top()` plots QTET of top-ranked samples for comparison against an experimental benchmark.
```{r}
plot_qte_top <- function(qtet_top, qtet_top0, bm, plot_titles, main_start = 1, 
                         ylim = c(-25000, 15000), col = NULL) {
  n <- length(qtet_top)
  for (i in 1:n) {
    main_title <- plot_titles[main_start + i - 1]
    mod <- qtet_top[[i]]
    mod2 <- qtet_top0[[i]]
    plot_qte(mod, mod2, bm, main = main_title, ylim = ylim, col = col)
    legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"),
           lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")
  }
}
```

**Arguments**

- `qtet_top`: A list of objects containing QTET estimates for top-ranked samples.
- `qtet_top0`: A list of objects containing unadjusted QTET estimates.
- `bm`: Experimental benchmark.
- `plot_titles`:  A character vector of titles for each plot, matching the order of elements in `qtet_top`. 
- `main_start`: Numeric value index specifying where to start selecting titles from `plot_titles` (default is 1).
- `ylim`: A numeric vector specifying the y-axis limits for the plots.
- `col`: Color specification.

`save_qtet()` generates and saves quantile treatment effect on the treated (QTET) plots for a given list of estimates.
```{r}
save_qtet <- function(plots, plot_titles = NULL, main_start = 1, 
                      ylim = c(-25000, 15000), col = NULL, prefix = "ldw") {
  dir.create("graphs/lalonde", showWarnings = FALSE, recursive = TRUE)
  n <- length(plots)
  for (i in seq_len(n)) {
    p <- plots[[i]]
    main_title <- if (is.null(plot_titles)) p$main else plot_titles[main_start + i - 1]
    clean_title <- gsub("[^a-zA-Z0-9]", "_", main_title)
    file_name <- sprintf("graphs/lalonde/%s_qtet_estimates_%s.pdf", prefix, clean_title)
    pdf(file = file_name, width = 7, height = 5)
    plot_qte(p$mod, p$mod0, p$bm, main = main_title, ylim = ylim, col = col)
    legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"),
           lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")
    dev.off()
  }
}
```

**Arguments**

- `plots`: A list of objects containing QTET estimates and corresponding unadjusted estimates.
- `plot_titles`: A character vector of titles for each plot, matching the order of elements in `plots`.
- `main_start`: Numeric value index specifying where to start selecting titles from `plot_titles` (default is 1).
- `ylim`: A numeric vector specifying the y-axis limits for the plots.
- `col`: Color specification.
- `prefix`: A string prefix for saved PDF filenames (default is "ldw").

`save_qte_top()` generates and saves quantile treatment effect on the treated (QTET) plots for a given list of estimates of top-ranked samples.
```{r}
save_qte_top <- function(qtet_top, qtet_top0, bm, plot_titles, main_start = 1,
                                ylim = c(-25000, 15000), col = NULL, prefix = "ldw_top") {
  n <- length(qtet_top)
  dir.create("graphs/lalonde", showWarnings = FALSE, recursive = TRUE)
  for (i in seq_len(n)) {
    mod <- qtet_top[[i]]
    mod2 <- qtet_top0[[i]]
    main_title <- plot_titles[main_start + i - 1]
    clean_title <- gsub("[^a-zA-Z0-9]", "_", main_title)
    file_name <- sprintf("graphs/lalonde/%s_qte_estimates_%s.pdf", prefix, clean_title)
    pdf(file = file_name, width = 7, height = 5)
    plot_qte(mod, mod2, bm, main = main_title, ylim = ylim, col = col)
    legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"),
           lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")
    dev.off()
  }
}
```

**Arguments**

- `qtet_top`: A list of objects containing QTET estimates for top-ranked samples.
- `qtet_top0`: A list of objects containing unadjusted QTET estimates.
- `plot_titles`: A character vector of titles for each plot, matching the order of elements in `qtet_top`.
- `main_start`: Numeric value index specifying where to start selecting titles from `plot_titles` (default is 1).
- `ylim`: A numeric vector specifying the y-axis limits for the plots.
- `col`: Color specification.
- `prefix`: A string prefix for saved PDF filenames (default is "ldw_top").

<br>

<a name="ow-ana"></a>

### Outcome weights analysis
`get_res_att()` estimates average treatment effects on the treated (ATT) using augmented inverse probability weighting with cross-fitting on a list of datasets.
```{r}
get_res_att <- function(dataset_list, Y, treat, covar,
                            estimator = "AIPW_ATT",
                            smoother = "honest_forest",
                            n_cf_folds = 5,
                            n_reps = 1) {
  lapply(dataset_list,
      function(data) {
        dml_with_smoother(
          Y = data[[Y]],
          D = data[[treat]],
          X = data[, covar, drop = FALSE],
          estimator = estimator,
          smoother = smoother,
          n_cf_folds = n_cf_folds,
          n_reps = n_reps
      )
    }
  )
}
```

**Arguments**

- `dataset_list`: A list of datasets. 
- `Y`: String specifying the name of the outcome variable.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `estimators`: String specifying the estimator to use (default is "AIPW_ATT").
- `smoother`:  String specifying the smoothing method (default is "honest_forest").
- `n_cf_folds`: Numeric value specifying the number of cross-fitting folds (default is 5).
- `n_reps`: Numeric value specifying the number of repetitions for robustness (default is 1).

`derive_ow()` extracts outcome weights from a list of estimation results.
```{r}
derive_ow <- function(results_list) {
lapply(results_list, function(res) get_outcome_weights(res))
}
```

**Argument**

- `results_list`: A list of objects obtained from `get_res_att()`.

`plot_ow()` plots the distribution of the outcome weights.
```{r}
plot_ow <- function(outcome_weights, plot_titles = NULL, breaks = 50, 
                    col = "#ff000080", xlab = "Outcome Weight", 
                    estimator = "AIPW-ATT") {
  N <- length(outcome_weights)
  for (i in seq_len(N)) {
    weights <- outcome_weights[[i]]$omega[estimator, ]
    main_title <- if (!is.null(plot_titles)) plot_titles[i] else paste("Dataset", i)
    hist(weights, breaks = breaks, main = main_title, xlab = xlab, col = col)
    mtext(paste("N =", length(weights)), side = 3, line = -1.5, cex = 0.8)
  }
  par(mfrow = c(1, 1))
}
```

**Arguments**

- `outcome_weights`: A list of outcome weight objects.
- `plot_titles`: A character vector of titles for each plot.
- `breaks`: Numeric value specifying the number of bins for the histogram plots (default is 50).
- `col`: Color specification (default is "#ff000080").
- `xlab`: String specifying the label for the x-axis (default is "Outcome Weight")
- `estimator`: String specifying the estimator to use (default is "AIPW-ATT").

`eval_ow()` computes summary statistics of outcome weights by treatment group for a list of datasets and returns a data frame for comparison.
```{r}
eval_ow <- function(outcome_weights, dataset_list, plot_titles = NULL, treat = "treat", estimator = "AIPW-ATT") {
  results <- lapply(seq_along(outcome_weights), function(i) {
    ow <- outcome_weights[[i]]$omega[estimator, ]
    treat <- dataset_list[[i]][[treat]]
    method <- if (!is.null(plot_titles)) plot_titles[i] else paste("Dataset", i)
    sum_treated <- sum(ow[treat == 1])
    sum_untreated <- sum(ow[treat == 0])
    data.frame(
      Method = method,
      Sum_Treated = sum_treated,
      Sum_Untreated = sum_untreated,
      stringsAsFactors = FALSE
    )
  })
  do.call(rbind, results)
}
```

**Arguments*

- `outcome_weights`: A list of outcome weight objects.
- `plot_titles`:  A character vector of titles for each plot.
- `treat_var`:  String specifying the treatment variable in each dataset (default is "treat").
- `estimator`: String specifying the estimator to use (default is "AIPW-ATT").

`save_ow()` saves the outcome weight histograms as PDF files, one per dataset, with optional custom titles and formatting.
```{r}
save_ow <- function(outcome_weights, plot_titles = NULL,
                    breaks = 50, col = "#ff000080", xlab = "Outcome Weight",
                    prefix = "model_a", estimand = "AIPW-ATT") {
  dir.create("graphs/lalonde", showWarnings = FALSE, recursive = TRUE)
  N <- length(outcome_weights)
  for (i in seq_len(N)) {
    file_name <- sprintf("graphs/lalonde/%s_outcomewt_%d.pdf", prefix, i)
    pdf(file = file_name, width = 8, height = 6)
    weights <- outcome_weights[[i]]$omega[estimand, ]
    main_title <- if (!is.null(plot_titles)) plot_titles[i] else paste("Dataset", i)
    hist(weights, breaks = breaks, main = main_title, xlab = xlab, col = col)
    mtext(paste("N =", length(weights)), side = 3, line = -1.5, cex = 0.8)
    dev.off()
  }
}
```

**Argument**

- `outcome_weights`: A list of outcome weight objects.
- `plot_titles`:  A character vector of titles for each plot.
- `breaks`: Numeric value specifying the number of bins for the histogram plots (default is 50).
- `col`: Color specification (default is "#ff000080").
- `xlab`: String specifying the label for the x-axis (default is "Outcome Weight")
- `prefix`: A string prefix for saved PDF filenames.
- `estimator`:  String specifying the estimator to use (default is "AIPW-ATT").

<a name="sens-ana"></a>

### Sensitivity analysis

`check_filter_datasets()` filters a list of datasets to retain only those with all required variables, no missing data, binary treatment, and sufficient baseline/covariate variation.
```{r}
check_filter_datasets <- function(datasets, Y, treat, covar, bm) {
  valid_datasets <- list()
  for (i in seq_along(datasets)) {
    data <- datasets[[i]]
    name <- names(datasets)[i]
    if (is.null(name) || name == "") name <- paste0("dataset_", i)  # provide a name if missing
    vars_needed <- c(Y, treat, covar, bm)
    if (!all(vars_needed %in% names(data))) { # check all variables exist
      message("Removed ", name, ": missing required variables") 
      next
    }
    sub <- data[, vars_needed, drop = FALSE]
    if (any(is.na(sub))) { # check no missing values
      message("Removed ", name, ": contains missing values") 
      next
    }
    tvals <- unique(sub[[treat]])  # check treatment binary
    if (!all(tvals %in% c(0, 1)) && !all(tvals %in% c(TRUE, FALSE))) {
      message("Removed ", name, ": treatment variable not binary")
      next
    }
    if (any(sapply(sub[bm], function(x) length(unique(x)) <= 1))) { # check baseline variables have variation
      message("Removed ", name, ": baseline variable(s) lack variation") 
      next
    }
    if (any(sapply(sub[covar], function(x) length(unique(x)) <= 1))) { # check covariates have variation
      message("Removed ", name, ": covariate variable(s) lack variation")
      next
    }
    valid_datasets[[name]] <- data # if all passed dataset is kept
  }
  return(valid_datasets)
}
```

**Arguments**

- `datasets`: A list of data frames containing data to be filtered.
- `Y`: The name of the outcome variable.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `bm`: A vector of baseline variable names.

`save_sensitivity_plots()` generates sensitivity analysis plots for a series of filtered datasets, saving each as a pdf file.
```{r}
save_sensitivity_plots <- function(filtered_datasets, Y, treat, covar, bm, plot_titles, prefix) {
  folder <- "graphs/lalonde"
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  for (i in seq_along(filtered_datasets)) {
    idx <- i
    file_name <- file.path(folder, paste0(prefix, "_sensitivity_", idx, ".pdf"))
    pdf(file_name, width = 8, height = 8)
    sens_ana(filtered_datasets[[i]], Y, treat, covar, bm, kd = 1:3)
    if (!missing(plot_titles) && length(plot_titles) >= idx) {
      title(main = plot_titles[idx])
    }
    dev.off()
 }
}
```

**Arguments**

- `filtered_datasets`: A list of filtered datasets.
- `Y`: The name of the outcome variable.
- `treat`: The name of the binary treatment indicator variable as string.
- `covar`: A vector of covariate names.
- `bm`: A vector of baseline variable names.
- `plot_titles`: A vector of titles for plots corresponding to datasets.
- `prefix`: A string prefix for saving output PDF files.

<div class="callout-tip">
Please refer to the accompanying paper for a detailed explanation of the metrics used in the analysis, including absolute standardized mean differences (SMDs), effective sample sizes (ESS), and the composite scoring approach employed to evaluate covariate balance and overlap across multiple methods. 
</div>