# LaLonde-Dehejia-Wahba (LDW) Data
LaLonde (1986) evaluated the effect of the National Supported Work Demonstration (NSW) program on both female and male participants. While the female participants were drawn from the Aid to Families with Dependent Children (AFDC) program, the male participants were drawn from three other groups: former drug addicts, ex-criminal offenders, and high-school dropouts. LaLonde relied on two main data sources:
(1) CPS-SSA-1, based on Westat’s Matched Current Population Survey–Social Security Administration File for individuals under 55 matching specific criteria;
(2) PSID-1, based on Panel Study of Income Dynamics for household heads under 55 from specific years who were not retired in 1975, adjusted for factors like employment status and poverty level, resulting in four additional comparison groups.

Dehejia and Wahba (1999) constructed a subset of LaLonde's original data, retaining 62% of observations and restricting the focus to male participants based on their 1974 earnings. As the dataset was formed entirely on pretreatment information like employment records or month of assignment, treatment allocation was effectively independent of all pre-treatment variables. Thus, they concluded that this subset, known as the LaLonde-Dehejia-Wahba (LDW) data, constitutes a valid experimental sample.

The analysis in section 2 and 3 builds on the LDW data and considers in total three samples: 
(1) LDW-Experimental, consisting of 185 treated individuals and 280 controls from the experimental data;
(2) LDW-CPS1, which pairs the same treated sample with 15,992 controls from CPS-SSA-1;
(3) LDW-PSID1, featuring the same treated sample with 2,490 controls from PSID-1; 
In section 4, the analysis applies the same set of statistical tools to analyze a fourth sample as an additional demonstration:
(4) LaLonde male samples (1986).
In the last section, the analysis applies a reduced set of statistical tools to analyze a fifth sample again as an additional demonstration:
(5) LaLonde female samples (2017).

This section (2) covers model A, which includes the outcome variable 1978 earnings (`re78`) and adjusts for a set of covariates: age, education, race (`black`, `hispanic`), marital status, high school dropouts, 1974 and 1975 earnings (`re74`, `re75`), and unemployment status in 1974 and 1975 (`u74`, `u75`). The model is defined by a regression formula for treatment assignment using these covariates. To improve covariate balance and overlap between treated and control groups, various methods are applied and are structured into five parts (matching, weighting, truncation, trimming and integrated methods). From these methods, the five best methods are determined based on a score defined by absolute standardized mean differences (SMD) and the effective sample size (ESS). The corresponding datasets are used to estimate the average treatment effect on the treated (ATT). Notably, the estimation incorporates the recently introduced augmented inverse probability weighting (AIPW) estimator, implemented via the `OutcomeWeights` R package. Subsequently, alternative estimands are considered, including the conditional average treatment effect for the treated (CATT) and the quantile treatment effect on the treated (QTET). After estimating these effects, outcome weights are analyzed to identify potential patterns in the contribution of individual observations to the ATT.  Placebo tests are then conducted using 1975 earnings (`re75`) as an outcome to assess potential biases and the validity of unconfoundedness assumptions. Finally, sensitivity analyses are performed to evaluate the robustness of treatment effect estimates to violations of these assumptions. 

## Set up
### Source functions and load data
```{r, message=FALSE, warning=FALSE}
setwd("C:/Users/lkreisel/workspace/uni-projects/Application-of-Outcome-Weights-Framework-for-DML-to-Lalonde-Study")
source("https://github.com/xuyiqing/lalonde/blob/main/tutorial/functions.R?raw=TRUE")
source("tutorial/functions.R")
#source("C:/Users/lkreisel/workspace/uni-projects/Application-of-Outcome-Weights-Framework-for-DML-to-Lalonde-Study/tutorial/functions.R")
# source functions (skip remote copy when offline)
# remote_url <- "https://raw.githubusercontent.com/xuyiqing/lalonde/main/tutorial/functions.R"
# tryCatch(
#         source(remote_url, local = FALSE),
#         error = function(e) message("Skipping remote helper script: ", conditionMessage(e))
# )
# source("tutorial/functions.R")
```

```{r, message=FALSE, warning=FALSE}
# load data
load("data/lalonde.RData")
```

```{r, message=FALSE, warning=FALSE}
# set seed
set.seed(42)
```

```{r, message=FALSE, warning=FALSE}
# enable plot display in VS Code
options(device = "windows")  # Opens plots in external window
```

### Inspect data

We begin the analysis with an overview of each dataset, where the dataset name, number of observations (rows), number of treated and control units, number of variables (columns), and variable names are reviewed.
```{r, message=FALSE, warning=FALSE}
# collect datasets in a list
data <- list(lalonde = lalonde, ldw_tr = ldw_tr, ldw_co = ldw_co, ldw_cps = ldw_cps, ldw_psid = ldw_psid)

# inspect key metrics of each dataset
summary_stats <- inspect_data(data)
datatable(summary_stats, caption = "Summary Statistics",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

### Load and preprocess data

Next, we augment the control groups in LDW-CPS1 and LDW-PSID1 by merging them with the experimental controls from LDW-Experimental. This approach follows the methodology laid out by @imbens2024. These expanded datasets are used solely for comparative purposes, while all primary analyses rely on the original LDW‑CPS1 and LDW‑PSID1 data.
```{r, message=FALSE, warning=FALSE}
# assigns 1 to the experimental controls
ldw_co$treat <- 1

# merge experimental data with CPS1 data
ldw_cps_plus <- rbind.data.frame(
  ldw_co,    # experimental controls 
  ldw_cps    # CPS1 data 
)

# merge experimental data with PSID1 data
ldw_psid_plus <- rbind.data.frame(
  ldw_co,    # experimental controls 
  ldw_psid   # PSID1 data 
)

datasets <- list(ldw_cps_plus  = ldw_cps_plus, ldw_psid_plus = ldw_psid_plus)

# inspect each dataset
summary_stats_plus <- inspect_data(datasets)
datatable(summary_stats_plus, caption = "Summary Statistics",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

## Model A

Finally, we define Model A as the baseline specification underlying the analysis.
```{r, message=FALSE, warning=FALSE}
# define variables
Y <- "re78" 
treat <- "treat" 
covar <- c("age", "education", "black", "hispanic", "married", 
           "nodegree", "re74", "re75", "u74", "u75") #re74 included
```

## Assessing overlap and covariate balance
### Overlap
To identify the average causal effect under unconfoundedness (that asserts that the treatment assignment is independent of the potential outcomes), it must be ensured that one can estimate the average effect at every value for the covariates. Thus, overlaps between the treated and untreated units (meaning that for every combination of covariates, there are both treated and untreated units in the dataset) are required. To assess overlap in the propensity scores and visualize the results using histograms of their log-odds, we use the `assess_overlap()` function of @imbens2024.

<div class="callout-note">
In a properly balanced experimental design, the distributions of the treatment group (red) and the control group (gray) would show considerable coverage.
</div>
```{r, fig.cap='FIGUREA1. SubfigureA:LDW. SubfigureB:LDW-CPS1. SubfigureC:LDW-PSID1.',out.width='100%', fig.asp=0.5}
ldw.ps <- assess_overlap(data = ldw, treat = treat, cov = covar)
ldw_cps.ps <- assess_overlap(data = ldw_cps, treat = treat, cov = covar) 
ldw_psid.ps <- assess_overlap(data = ldw_psid, treat = treat, cov = covar) 
```

As anticipated, LDW-Experimental exhibits an almost perfect overlap. In contrast, the observational datasets LDW-CPS1 and LDW-PSID1 show weak overlap. Notably, many treated units have propensity scores outside the range of the controls, while a large share of control units have propensity scores concentrated at very low log-odds.

Next, the overlap of the expanded observational datasets is examined.
```{r, fig.cap='FIGUREA1. SubfigureD:LDW-CPS1-PLUS. SubfigureE:LDW-PSID1-PLUS.', out.width='80%', fig.asp=1, fig.align='center'}
ldw_cps_plus.ps <- assess_overlap(data = ldw_cps_plus, treat = treat, cov = covar) 
ldw_psid_plus.ps <- assess_overlap(data = ldw_psid_plus, treat = treat, cov = covar)
```

As expected, the datasets LDW-CPS1-PLUS and LDW-PSID1-PLUS show an improved overlap compared to their non-plus states.

For the following analysis, we set up a model formula.
```{r, message=FALSE, warning=FALSE}
# set model formula
model <- as.formula(paste(treat, "~", paste(covar, collapse = " + ")))
```

### Covariate balance

To make treatment and control groups more comparable, thus mimicking a randomized experiment, initial covariate balance is assessed to determine the degree to which treatment and control groups differ on observed characteristics. Therefore, we employ visual summaries using `love.plot()` by @imbens2024, which depicts standardized mean differences across covariates before and after adjustment.
```{r, warning=FALSE, message=FALSE, fig.cap='FIGUREA2. SubfigureA:LDW-CPS1. SubfigureB:LDW-CPS1-PLUS. SubfigureC:LDW-PSID1. SubfigureD:LDW-PSID1-PLUS.', out.width='80%', fig.asp=1, fig.align='center'}
# plot balance
love.plot(ldw, ldw_cps, treat, covar = covar, title = "LDW-CPS1")
love.plot(ldw, ldw_psid, treat, covar = covar, title = "LDW-PSID1")
love.plot(ldw, ldw_cps_plus, treat, covar = covar, title = "LDW-CPS1-PLUS")
love.plot(ldw, ldw_psid_plus, treat, covar = covar, title = "LDW-PSID1-PLUS")
```

Neither LDW-CPS1-PLUS nor LDW-PSID1-PLUS shows consistently improved covariate balance compared to their non-plus counterparts based on raw standardized mean differences. Although some specific covariates improve slightly, most show increased imbalance.

<div class="callout-note">
For the subsequent analysis aimed at enhancing covariate balance and overlap, only the two datasets LDW-CPS1 and LDW-PSID1 are used. The LDW-Experimental dataset is excluded from these steps, as randomization already ensures adequate covariate balance and overlap. 
</div>

## Improving overlap
### Single methods
#### Trimming 

The purpose of trimming is to remove units whose propensity scores are too dissimilar from the opposite group, thereby improving covariate balance between treated and control groups. Below, we employ several trimming methods in line with the approaches proposed by @Stuermer_2021 and @imbens2024.

1) Propensity score threshold trimming 
```{r, message=FALSE, warning=FALSE}
# apply trimming with threshold 0.9
ldw_cps.ps_trim <- ps_trim(ldw_cps.ps, threshold = 0.9)
ldw_psid.ps_trim <- ps_trim(ldw_psid.ps, threshold = 0.9)

# re-estimate propensity scores on trimmed data 
ldw_cps.ps_trim <- ps_estimate(data = ldw_cps.ps_trim, treat = treat, cov = covar)
ldw_psid.ps_trim <- ps_estimate(data = ldw_psid.ps_trim, treat = treat, cov = covar)
```

2) Common range trimming
```{r, message=FALSE, warning=FALSE}
# trim observations outside the common support region of propensity scores
ldw_cps.ps_common   <- common_range_trim(ldw_cps.ps)
ldw_psid.ps_common  <- common_range_trim(ldw_psid.ps)

# re-estimate propensity scores on trimmed data 
ldw_cps.ps_common <- ps_estimate(data = ldw_cps.ps_common, treat = treat, cov = covar)
ldw_psid.ps_common <- ps_estimate(data = ldw_psid.ps_common, treat = treat, cov = covar)
```

3) Crump trimming 
```{r, message=FALSE, warning=FALSE}
# trim observations with propensity scores outside an interval
ldw_cps.ps_crump  <- crump_trim(ldw_cps.ps, lower = 0.05, upper = 0.95)
ldw_psid.ps_crump <- crump_trim(ldw_psid.ps, lower = 0.05, upper = 0.95)

# re-estimate propensity scores on trimmed data 
ldw_cps.ps_crump <- ps_estimate(data = ldw_cps.ps_crump, treat = treat, cov = covar)
ldw_psid.ps_crump <- ps_estimate(data = ldw_psid.ps_crump, treat = treat, cov = covar)
```

4) Stuermer trimming
```{r, message=FALSE, warning=FALSE}
# trim observations based on propensity score quantiles separately for treated and control
ldw_cps.ps_stuermer  <- stuermer_trim(ldw_cps.ps)
ldw_psid.ps_stuermer <- stuermer_trim(ldw_psid.ps)

# re-estimate propensity scores on trimmed data 
ldw_cps.ps_stuermer <- ps_estimate(data = ldw_cps.ps_stuermer, treat = treat, cov = covar)
ldw_psid.ps_stuermer <- ps_estimate(data = ldw_psid.ps_stuermer, treat = treat, cov = covar)
```

5) Walker trimming
```{r, message=FALSE, warning=FALSE}
# trim observations based on preference scores that adjust for treatment prevalence using logit transformations
ldw_cps.ps_walker  <- walker_trim(ldw_cps.ps)
ldw_psid.ps_walker <- walker_trim(ldw_psid.ps)

# re-estimate propensity scores on trimmed data 
ldw_cps.ps_walker <- ps_estimate(data = ldw_cps.ps_walker, treat = treat, cov = covar)
ldw_psid.ps_walker <- ps_estimate(data = ldw_psid.ps_walker, treat = treat, cov = covar)
```

### Integrated methods
#### Trimming and matching 

The purpose of matching is to create comparable treated and control groups by pairing units with similar covariate profiles, thereby reducing covariate imbalance and confounding bias. We apply several matching methods as outlined by @Greifer_3_2025 in the following. 

<div class="callout-note">
For comparison, combinations of trimming and matching methods are applied in line with the tutorial by @imbens2024, before being applied to the initial datasets.
</div>

##### Extended datasets 
(Similar to tutorial by @imbens2024)
```{r, message=FALSE, warning=FALSE}
ldw_cps_trim <- ps_trim(ldw_cps_plus.ps, threshold = 0.9)
ldw_psid_trim <- ps_trim(ldw_psid_plus.ps, threshold = 0.8)

# excluding the experimental controls
ldw_cps.trim_match <- subset(ldw_cps_trim, sample %in% c(1,3) & ps_assoverlap)
ldw_psid.trim_match <- subset(ldw_psid_trim, sample %in% c(1,4) & ps_assoverlap)

# re-estimate propensity scores and employ 1:1 matching
ldw_cps.trim_match <- psmatch(data = ldw_cps.trim_match, Y = "re78", treat = "treat", cov = covar)
ldw_psid.trim_match <- psmatch(data = ldw_psid.trim_match, Y = "re78", treat = "treat", cov = covar)

# further subset data and re-assign treat variable 
ldw_trim_cps <- subset(ldw_cps_trim, sample %in% c(1,2) & ps_assoverlap <= 0.9)
ldw_trim_cps$treat[which(ldw_trim_cps$sample == 2)] <- 0

ldw_trim_psid <- subset(ldw_psid_trim, sample %in% c(1,2) & ps_assoverlap <= 0.8)
ldw_trim_psid$treat[which(ldw_trim_psid$sample == 2)] <- 0
```

##### Initial datasets
```{r, message=FALSE, warning=FALSE}
all_trim.cps  <- list(
  ps_threshold = ldw_cps.ps_trim, 
  common_range = ldw_cps.ps_common, 
  stuermer = ldw_cps.ps_stuermer, 
  walker = ldw_cps.ps_walker, 
  crump = ldw_cps.ps_crump)

all_trim.psid  <- list(
  ps_threshold = ldw_psid.ps_trim, 
  common_range = ldw_psid.ps_common, 
  stuermer = ldw_psid.ps_stuermer, 
  walker = ldw_psid.ps_walker, 
  crump = ldw_psid.ps_crump)
```

###### Distance matching
1) 1:1 Nearest neighbor matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform nearest neighbor matching with k=1, logistic propensity score and replacement
nn_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "nearest", distance = "glm", link = "logit", replace = TRUE)
nn_trim_comb.psid <- attach_matchit(model, data_list = all_trim.psid, method = "nearest", distance = "glm", link = "logit", replace = TRUE)
```

2) k:1 matching (k=2) with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform nearest neighbor matching with k=2, logistic propensity score and replacement
k<-2
k2_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "nearest", distance = "glm", link = "logit", ratio = k, replace = TRUE)
k2_trim_comb.psid <- attach_matchit(model, data_list = all_trim.psid, method = "nearest", distance = "glm", link = "logit", ratio = k, replace = TRUE)
```

3) k:1 matching (k=3) with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform nearest neighbor matching with k=3, logistic propensity score and replacement
k<-3
k3_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "nearest", distance = "glm", link = "logit", ratio = k, replace = TRUE)
k3_trim_comb.psid <- attach_matchit(model, data_list = all_trim.psid, method = "nearest", distance = "glm", link = "logit", ratio = k, replace = TRUE)
```

4) Caliper matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform nearest neighbor matching with a caliper of 0.1 on the logistic propensity score 
caliper_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "nearest", distance = "glm", link = "logit", caliper = 0.1, replace = TRUE)
caliper_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "nearest", distance = "glm", link = "logit", caliper = 0.1, replace = TRUE)
```

5) Common support restriction matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform nearest neighbor matching with exclusion of units outside common support
cs_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "nearest", distance = "glm", link = "logit", discard = "both", reestimate = TRUE, replace = TRUE)
cs_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "nearest", distance = "glm", link = "logit", discard = "both", reestimate = TRUE, replace = TRUE)
```

6) Mahalanobis distance matching (mahvars) with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform nearest neighbor matching using mahalanobis distance
mahvars_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "nearest", distance = "mahalanobis", replace = FALSE)
mahvars_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "nearest", distance = "mahalanobis", replace = FALSE)
```

7) Optimal pair matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform optimal pair matching that minimizes total within-pair distance on propensity scores
optimal_pair_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "optimal", distance = "glm", link = "logit")
optimal_pair_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "optimal", distance = "glm", link = "logit")
```

8) Optimal full matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform optimal full matching allowing sets with varying ratios of treated to controls and minimizing a global distance criterion
optimal_full_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "full", distance = "glm", link = "logit")
optimal_full_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "full", distance = "glm", link = "logit")
```

9) Generalized full matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform generalized full matching using fast approximation and producing matched sets with flexible treated/control ratios 
general_full_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "quick", distance = "glm", link = "logit")
general_full_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "quick", distance = "glm", link = "logit")
```

10) Genetic matching with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump trimming
```{r, message=FALSE, warning=FALSE}
# perform genetic matching 
genetic_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "genetic", distance = "glm", link = "logit", replace = TRUE, pop.size = 100)
genetic_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "genetic", distance = "glm", link = "logit", replace = TRUE, pop.size = 100)
```

###### Stratum matching
11) Exact matching (exact)
```{r, message=FALSE, warning=FALSE}
# match units exactly by raw covariate profiles 
exact_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "exact")
exact_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "exact")
```

12) Coarsened matching (cem)
```{r, message=FALSE, warning=FALSE}
# match units exactly within coarse strata 
cem_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "cem")
cem_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "cem")
```

13) Subclassification 
```{r, message=FALSE, warning=FALSE}
# partition sample into fixed number of bins based on propensity score 
subcl_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "subclass", subclass = 3)
subcl_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "subclass", subclass = 3)
```

###### Pure subset selection
14) Cardinality profile matching
```{r, message=FALSE, warning=FALSE}
# select largest balanced subsample meeting covariate balance tolerances and a fixed ratio of controls to treated units
card_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "cardinality", tols = 0.1, ratio = 1, time = 1200)
card_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "cardinality", tols = 0.1, ratio = 1, time = 1200)
```

15) Profile matching
```{r, message=FALSE, warning=FALSE}
# select the largest control group subset balanced to the treated group (for ATT), leaving the treated group intact
profile_trim_comb.cps <- attach_matchit(model, data_list = all_trim.cps, method = "cardinality", estimand = "ATT", tols = 0.1, ratio = NA, time = 1200)
profile_trim_comb.psid <- attach_matchit(model,data_list = all_trim.psid, method = "cardinality", estimand = "ATT", tols = 0.1, ratio = NA, time = 1200)
```

## Assessing methods

We now systematically reassess all methods described above by evaluating the overlap coefficient (OVL) and the absolute standardized mean differences (ASMD). Specifically, we compute the ASMD and OVL for each approach and compute a composite score to determine the best performing methods.

### Single methods
#### Trimming
##### SMD
```{r, message=FALSE, warning=FALSE}
# compute absolute SMD
smd_trim.cps <- compute_abs_smd_trim(all_trim.cps, "treat", covar)
smd_trim.psid <- compute_abs_smd_trim(all_trim.psid, "treat", covar)
```

##### OVL
```{r, message=FALSE, warning=FALSE}
# compute overlap coefficients
ovl_trim.cps <- compute_ovl_trim(all_trim.cps, "ps_assoverlap", "treat")
ovl_trim.psid <- compute_ovl_trim(all_trim.psid, "ps_assoverlap", "treat")
```

### Integrated methods
#### Trimming and matching
##### Extended datasets
(Similar to tutorial by @imbens2024)
```{r, message=FALSE, warning=FALSE}
# list all trimmed and matched samples
trim_match_comb.cps_plus <- list(ps_threshold_match = ldw_cps.trim_match)
trim_match_comb.psid_plus <- list(ps_threshold_match = ldw_psid.trim_match)
```

###### SMD
```{r, message=FALSE, warning=FALSE}
# compute absolute SMD
smd_trim_match_comb.cps_plus <- compute_abs_smd_trim(trim_match_comb.cps_plus, "treat", covar)
smd_trim_match_comb.psid_plus <- compute_abs_smd_trim(trim_match_comb.psid_plus, "treat", covar)
```

###### OVL
```{r, message=FALSE, warning=FALSE}
# compute overlap coefficients
ovl_trim_match_comb.cps_plus <- compute_ovl_trim(trim_match_comb.cps_plus, "ps_assoverlap", "treat")
ovl_trim_match_comb.psid_plus <- compute_ovl_trim(trim_match_comb.psid_plus, "ps_assoverlap", "treat")
```

##### Initial datasets
```{r, message=FALSE, warning=FALSE}
# list all trimmed and matched samples
trim_match_comb.cps <- list(
  nn = nn_trim_comb.cps, 
  k2 = k2_trim_comb.cps,
  k3 = k3_trim_comb.cps, 
  caliper = caliper_trim_comb.cps,
  cs = cs_trim_comb.cps, 
  mahvars = mahvars_trim_comb.cps,
  optimal_pair = optimal_pair_trim_comb.cps,
  optimal_full = optimal_full_trim_comb.cps, 
  genetic = genetic_trim_comb.cps, 
  exact = exact_trim_comb.cps,
  cem = cem_trim_comb.cps,
  subcl = subcl_trim_comb.cps 
)

trim_match_comb.psid <- list(
  nn = nn_trim_comb.psid,
  k2 = k2_trim_comb.psid, 
  k3 = k3_trim_comb.psid, 
  caliper = caliper_trim_comb.psid, 
  cs = cs_trim_comb.psid,
  mahvars = mahvars_trim_comb.psid,
  optimal_pair = optimal_pair_trim_comb.psid, 
  optimal_full = optimal_full_trim_comb.psid, 
  genetic = genetic_trim_comb.psid,
  exact = exact_trim_comb.psid,
  cem = cem_trim_comb.psid,
  subcl = subcl_trim_comb.psid 
)
```

###### SMD
```{r, message=FALSE, warning=FALSE}
# compute absolute SMD
smd_trim_match_comb.cps <- compute_abs_smd_matchit(trim_match_comb.cps, all_trim.cps) 
smd_trim_match_comb.psid <- compute_abs_smd_matchit(trim_match_comb.psid, all_trim.psid)
```

###### OVL
```{r, message=FALSE, warning=FALSE}
# compute overlap coefficients
ovl_trim_match_comb.cps <- compute_ovl_matchit(trim_match_comb.cps, all_trim.cps, ps = "ps_assoverlap", treat = "treat", covar = covar)
ovl_trim_match_comb.psid <- compute_ovl_matchit(trim_match_comb.psid, all_trim.psid, ps = "ps_assoverlap", treat = "treat", covar = covar)
```

## Identifying best methods

### Ranking
To identify the top five methods for each observational dataset, we first combine for each dataset all results of the absolute SMD and OVL into a single data frame. This allows for a comprehensive comparison across all methods.

<div class="callout-note">
Only results based on the non-plus datasets are included in the identification of the best methods, as the objective is to identify the overall best-performing methods for the original observational samples. 
</div>
```{r, message=FALSE, warning=FALSE}
# combine all results
all_cps <- combine_results("cps")
all_psid <- combine_results("psid") 
```

```{r, message=FALSE, warning=FALSE}
# save results
save_csv(all_cps, "ldw_model_a_cps1_all_results")
save_csv(all_psid, "ldw_model_a_psid1_all_results")
```

Next, each method is evaluated according to its OVL value. The top five methods are then selected based on the highest OVL values, with higher OVL values indicating better overlap between treated and control groups.
```{r, message=FALSE, warning=FALSE}
# rank comparatively
ranked_cps  <- assess_methods(all_cps)
ranked_psid <- assess_methods(all_psid)

# get top 5 methods
top5_methods.cps <- get_top_methods(ranked_cps, top_n = 5)
top5_methods.psid <- get_top_methods(ranked_psid, top_n = 5)

# rerank top 5 methods
top5_methods.cps <- rerank_methods(top5_methods.cps, ranked_cps)
top5_methods.psid <- rerank_methods(top5_methods.psid, ranked_psid)
```

```{r, message=FALSE, warning=FALSE}
# print results
top5_methods_df.cps <- ranked_cps %>% arrange(desc(OVL)) %>% head(5)
top5_methods_df.psid <- ranked_psid %>% arrange(desc(OVL)) %>% head(5)

datatable(top5_methods_df.cps, caption = "Top 5 Methods for CPS1",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
datatable(top5_methods_df.psid, caption = "Top 5 Methods for PSID1",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

```{r, message=FALSE, warning=FALSE}
# save results 
save_csv(top5_methods.cps, "ldw_model_a_cps1_top5_methods")
save_csv(top5_methods.psid, "ldw_model_a_psid1_top5_methods")
```

### Dataset construction
For the subsequent estimation, we need to construct datasets of the top‑ranked methods. Therefore, we match all method names back to their corresponding datasets, objects or vectors and construct corresponding datasets. 
```{r, message=FALSE, warning=FALSE}
trim_only_cps <- list(
        ps_threshold = ldw_cps.ps_trim,
        common_range = ldw_cps.ps_common,
        stuermer = ldw_cps.ps_stuermer,
        walker = ldw_cps.ps_walker,
        crump = ldw_cps.ps_crump
)

match_lookup_cps <- c(
        wrap_match_entries(nn_trim_comb.cps, all_trim.cps, "nn"),
        wrap_match_entries(k2_trim_comb.cps, all_trim.cps, "k2"),
        wrap_match_entries(k3_trim_comb.cps, all_trim.cps, "k3"),
        wrap_match_entries(caliper_trim_comb.cps, all_trim.cps, "caliper"),
        wrap_match_entries(cs_trim_comb.cps, all_trim.cps, "cs"),
        wrap_match_entries(mahvars_trim_comb.cps, all_trim.cps, "mahvars"),
        wrap_match_entries(optimal_pair_trim_comb.cps, all_trim.cps, "optimal_pair"),
        wrap_match_entries(optimal_full_trim_comb.cps, all_trim.cps, "optimal_full"),
        wrap_match_entries(genetic_trim_comb.cps, all_trim.cps, "genetic"),
        wrap_match_entries(exact_trim_comb.cps, all_trim.cps, "exact"),
        wrap_match_entries(cem_trim_comb.cps, all_trim.cps, "cem"),
        wrap_match_entries(subcl_trim_comb.cps, all_trim.cps, "subcl")
)

trim_only_psid <- list(
        ps_threshold = ldw_psid.ps_trim,
        common_range = ldw_psid.ps_common,
        stuermer = ldw_psid.ps_stuermer,
        walker = ldw_psid.ps_walker,
        crump = ldw_psid.ps_crump
)

match_lookup_psid <- c(
        wrap_match_entries(nn_trim_comb.psid, all_trim.psid, "nn"),
        wrap_match_entries(k2_trim_comb.psid, all_trim.psid, "k2"),
        wrap_match_entries(k3_trim_comb.psid, all_trim.psid, "k3"),
        wrap_match_entries(caliper_trim_comb.psid, all_trim.psid, "caliper"),
        wrap_match_entries(cs_trim_comb.psid, all_trim.psid, "cs"),
        wrap_match_entries(mahvars_trim_comb.psid, all_trim.psid, "mahvars"),
        wrap_match_entries(optimal_pair_trim_comb.psid, all_trim.psid, "optimal_pair"),
        wrap_match_entries(optimal_full_trim_comb.psid, all_trim.psid, "optimal_full"),
        wrap_match_entries(genetic_trim_comb.psid, all_trim.psid, "genetic"),
        wrap_match_entries(exact_trim_comb.psid, all_trim.psid, "exact"),
        wrap_match_entries(cem_trim_comb.psid, all_trim.psid, "cem"),
        wrap_match_entries(subcl_trim_comb.psid, all_trim.psid, "subcl")
)

list_cps <- c(trim_only_cps, match_lookup_cps)
list_psid <- c(trim_only_psid, match_lookup_psid)
```

```{r, message=FALSE, warning=FALSE}
# create datasets corresponding to the top 5 methods for each dataset
top5_datasets.cps <- create_top5_datasets(list_cps, top5_methods.cps)
top5_datasets.psid <- create_top5_datasets(list_psid, top5_methods.psid)

# extract top5 objects for bal.plot()
top5_objects.cps <- lapply(top5_methods.cps, function(d) list_cps[[d]])
top5_objects.psid <- lapply(top5_methods.psid, function(d) list_psid[[d]])
```

```{r, message=FALSE, warning=FALSE}
# save datasets into .RData files
save_top5_datasets(list_cps, top5_methods.cps, prefix = "ldw_model_a_cps1")
save_top5_datasets(list_psid, top5_methods.psid, prefix = "ldw_model_a_psid1")
```

## Estimating
### Average treatment effect on the treated (ATT)

Next, we estimate the average treatment effect on the treated (ATT) using the LDW-Experimental sample, the top‑ranked observational samples (LDW‑CPS1 and LDW‑PSID1), and, for comparison, the trimmed and matched plus‑datasets. We employ a broad set of estimators, including difference-in-means, regression, the Oaxaca-Blinder estimator, generalized random forests (GRF) as an outcome model, 1:5 nearest neighbor matching with bias correction, inverse probability weighting (IPW) with GRF-based propensity scores, covariate balancing propensity score (CBPS), entropy balancing, double/debiased machine learning using elastic net, and augmented inverse probability weighting (AIPW) via GRF. We also incorporate the recently introduced estimand-specific AIPW estimator implemented through an expansion of the `OutcomeWeights` R package by @OW_Package.

We utilize the `estimate_all()` and `plot_coef()` functions as defined by @imbens2024.
```{r, message=FALSE, warning=FALSE}
# estimate ATT
out1 <- estimate_all(ldw, "re78", "treat", covar)
out2 <- estimate_all(ldw_cps, "re78", "treat", covar)
out3 <- estimate_all(ldw_psid, "re78", "treat", covar)

out.cps <- lapply(top5_datasets.cps, function(d) estimate_all(d, "re78", "treat", covar))
out.psid <- lapply(top5_datasets.psid, function(d) estimate_all(d, "re78", "treat", covar))

out4 <- out.cps[[1]]
out5 <- out.cps[[2]]
out6 <- out.cps[[3]]
out7 <- out.cps[[4]]
out8 <- out.cps[[5]]

out9 <- out.psid[[1]]
out10 <- out.psid[[2]]
out11 <- out.psid[[3]]
out12 <- out.psid[[4]]
out13 <- out.psid[[5]]
```

```{r, message=FALSE, warning=FALSE}
# estimate ATT
load("data/trimmed.RData")
out14 <- estimate_all(ldw_trim_cps, "re78", "treat", covar)
out15 <- estimate_all(ldw_trim_psid, "re78", "treat", covar)
out16 <- estimate_all(ldw_cps_trim, "re78", "treat", covar)
out17 <- estimate_all(ldw_psid_trim, "re78", "treat", covar)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA10. SubfigureA:LDW-Experimental. SubfigureB:LDW-CPS1. SubfigureC:LDW-PSID1. SubfigureD-H:Top-LDW-CPS1. SubfigureI-M:Top-LDW-PSID1. SubfigureN:LDW-CPS1-PLUS. SubfigureO:LDW-PSID1-PLUS. ATT Estimates Model A Given Unconfoundedness using LDW Samples"}
par(mfrow = c(4, 1), mar = c(4, 4, 2, 2))

# experimental benchmarks
band.exp <- out1[1, 3:4]
est.exp  <- out1[1, 1]

# plot results
plot_coef(out1,  band = band.exp, line = est.exp, 
          ylim = c(-15500, 5500), main = "(A) LDW-Experimental")

plot_coef(out2,  band = band.exp, line = est.exp, 
          ylim = c(-15500, 5500), main = "(B) LDW-CPS1")

plot_coef(out3,  band = band.exp, line = est.exp,
          ylim = c(-15500, 5500), main = "(C) LDW-PSID1")

for (i in seq_along(out.cps)) {
  this_title <- paste0("(", LETTERS[i+3], ") Top CPS1: ", top5_methods.cps[i])
  plot_coef(out.cps[[i]], band = band.exp, line = est.exp,
            ylim = c(-15500, 5500), main = this_title)
}

for (i in seq_along(out.psid)) {
  this_title <- paste0("(", LETTERS[i+8], ") Top PSID1: ", top5_methods.psid[i])
  plot_coef(out.psid[[i]], band = band.exp, line = est.exp,
            ylim = c(-15500, 5500), main = this_title)
}

# nonexperimtenal benchmarks
band.cps_plus <- out14[1, 3:4]
est.cps_plus <- out14[1, 1]
band.psid_plus <- out15[1, 3:4]
est.psid_plus <- out15[1, 1]

# plot results
plot_coef(out16, band = band.cps_plus, line = est.cps_plus, 
          ylim = c(-15500, 5500), main = "(N) Trimmed LDW-CPS1-PLUS")


plot_coef(out17, band = band.psid_plus, line = est.psid_plus, 
          ylim = c(-15500, 5500), main = "(O) Trimmed LDW-PSID1-PLUS")
```

```{r, message=FALSE, warning=FALSE}
# save results
save_att_panels(
  out_list = list(out1, out2, out3),
  plot_titles = c("(A) LDW-Experimental", "(B) LDW-CPS1", "(C) LDW-PSID1"),
  band_list = list(band.exp, band.exp, band.exp),
  est_list  = list(est.exp, est.exp, est.exp),
  prefix = "ldw_model_a_est_exp"
)

save_att_panels(
  out_list = out.cps,
  plot_titles = paste0("(", LETTERS[4:8], ") Top CPS1: ", top5_methods.cps),
  band_list = replicate(length(out.cps), band.exp, simplify = FALSE),
  est_list  = replicate(length(out.cps), est.exp, simplify = FALSE),
  prefix = "ldw_model_a_est_top_cps"
)

save_att_panels(
  out_list = out.psid,
  plot_titles = paste0("(", LETTERS[9:13], ") Top PSID1: ", top5_methods.psid),
  band_list = replicate(length(out.psid), band.exp, simplify = FALSE),
  est_list  = replicate(length(out.psid), est.exp, simplify = FALSE),
  prefix = "ldw_model_a_est_top_psid"
)

save_att_panels(
  out_list = list(out16, out17),
  plot_titles = c("(N) Trimmed LDW-CPS1-PLUS", "(O) Trimmed LDW-PSID1-PLUS"),
  band_list = list(band.cps_plus, band.psid_plus),
  est_list  = list(est.cps_plus, est.psid_plus),
  prefix = "ldw_model_a_est_plus"
)
```

The above figures show the ATT estimates and their 95% confidence intervals for fifteen samples: LDW-Experimental, LDW-CPS1, LDW-PSID1, a series of top-ranked subsamples of both LDW-CPS1 and LDW-PSID1 based on various matching, weighting, truncation and trimming criteria and trimmed and matched versions of the LDW-CPS1 and LDW-PSID1 samples (similar to @imbens2024).

Figure (A) presents the benchmark from the experimental sample (LDW-Experimental), serving as a reference for bias and variance assessment of observational samples. Figures (B) and (C) show results for the original observational samples, LDW-CPS1 and LDW-PSID1, while figures (D) through (H) display results for CPS1-based subsamples constructed with the top-ranked methods. Analogously, figures (I) through (M) summarize results for the corresponding PSID1-based subsamples under parallel rules. Figures (N) and (O) present results for the trimmed and matched versions, replicating the tutorial results of Imbens & Xu (2024). 

Across the LDW-CPS1 dataset and its top-ranked subsamples, all estimators generally produce ATT estimates that closely cluster around the experimental benchmark. However, the `ebal_weight` sample exhibit somewhat larger deviations. While most estimates are positive, the `ebal_weight` subsample results include one negative ATT estimates.

In contrast, PSID1-based subsamples frequently exhibit greater dispersion and substantially higher standard errors compared to LDW-CPS1 samples. While most ATT estimates are positive, all methods exhibit some negative ATT estimators across various estimators. 
```{r, message=FALSE, warning=FALSE}
# get all outputs
all_outs <- c(list(out1, out2, out3), out.cps, out.psid, list(out16, out17))

# get plot titles
all_plot_titles <- c("(A) LDW-Experimental", "(B) LDW-CPS1", "(C) LDW-PSID1",
                      paste0("(", LETTERS[4:8], ") Top CPS1: ", top5_methods.cps),
                      paste0("(", LETTERS[9:13], ") Top PSID1: ", top5_methods.psid),
                      "(N) Trimmed LDW-CPS1-PLUS", "(O) Trimmed LDW-PSID1-PLUS")

# evaluate results
all_summaries <- lapply(all_outs, eval_att)
att_summary <- do.call(rbind, all_summaries)
rownames(att_summary) <- all_plot_titles

# print results
datatable(att_summary, caption = "ATT Summary Statistics",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

The ATT results are presented in the table below:
```{r, message=FALSE, warning=FALSE}
# create result matrix
result_mat <- create_matrix_results(all_outs, all_plot_titles)

# render formatted table output
datatable(result_mat, caption = "ATT Estimates and SEs",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

<div class="callout-note">
The first value in each cell represents the point estimate from the respective estimator and sample. The number in brackets directly following it is the robust standard error (SE) associated with that estimate, indicating the uncertainty or variability around the point estimate.
</div>

The tabulated results confirm visual patterns: Column (A) reports the estimates for the LDW-Experimental sample, column (B) for the LDW-CPS1 sample, and column (C) for the LDW-PSID1 sample. Columns (D)-(O) summarize the trimmed and top-ranked sample results for both LDW-CPS1 and LDW-PSID1. 

For most LDW-CPS1-based samples, estimates remain not only positive but also comparatively close to the experimental benchmark, with modest variance inflation. In contrast, the LDW-PSID1-based estimates exhibit greater variability, more frequently negative values, and larger standard errors, reflecting an increased challenge of achieving balance and overlap in this observational dataset.

Overall, figures and table jointly demonstrate that ATT estimates from observational samples tend to have larger standard errors compared to the experimental sample, reflecting greater statistical uncertainty in non-experimental causal effect estimation, and that certain criteria can bring observational estimates closer to the experimental benchmark. Nevertheless, significant estimator-dependent variability and sensitivity to sample construction persist.
```{r, message=FALSE, warning=FALSE}
# save results 
save_csv(result_mat, "ldw_att_est_model_a")
```

<div class="callout-tip">
Improved overlap and covariate balance generally leads to estimates closer to the experimental benchmark, but often at the cost of increased standard errors. This trade-off highlights the need to carefully balance bias reduction against precision when selecting methods for robust causal inference.
</div>

Next, we explore alternative estimands such as heterogeneous treatment effects and quantile treatment effects, which may provide deeper insights into the validity of the unconfoundedness assumption.

### Conditional average treatment effect on the treated (CATT)

CATT enables analysis of treatment effect heterogeneity across covariate-defined subpopulations. We estimate the CATT for the same samples previously considered in the ATT estimation by applying the Augmented Inverse Probability Weighting Generalized Random Forest (AIPW-GRF) method via the `catt()` function from Imbens and Xu (2024).
```{r, message=FALSE, warning=FALSE}
# estimate CATT
catt.ldw <- catt(ldw, Y, treat, covar)
catt.cps <- catt(ldw_cps, Y, treat, covar)
catt.psid <- catt(ldw_psid, Y, treat, covar)
catt.top5_cps <- lapply(top5_datasets.cps, function(d) catt(d, Y, treat, covar))
catt.top5_psid <- lapply(top5_datasets.psid, function(d) catt(d, Y, treat, covar))
```

```{r, message=FALSE, warning=FALSE}
# similar to Imbens & Xu (2024) 
catt.trim_cps <- catt(ldw_trim_cps, Y, treat, covar) 
catt.trim_psid <- catt(ldw_trim_psid, Y, treat, covar) 
catt.cps_trim <- catt(ldw_cps_trim, Y, treat, covar)  
catt.psid_trim <- catt(ldw_psid_trim, Y, treat, covar) 
```

Then, we employ a modified version of the function `plot_catt()` from @imbens2024 to visualize the results by plotting the estimated CATTs for all samples at the covariate values of each treated unit against their corresponding experimental benchmarks. 

<div class="callout-note">
Gray dots represent pairs of CATT estimates at covariate values of treated units, while the red crosses indicate pairs of estimated ATTs.
</div>
```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA11. SubfigureB:LDW-CPS1. SubfigureC:LDW-PSID1. CATT Estimates Model A using LDW Data: Experimental vs. Nonexperimental"}
# plot results
par(mfrow = c(2,2)) 
par(cex.main = 0.9)

plot_catt(
  catt1 = catt.ldw$catt,
  catt2 = catt.cps$catt,
  att1  = catt.ldw$att[1],
  att2  = catt.cps$att[1],
  xlab  = "CATT (Experimental)",
  ylab  = "CATT (CPS1)",
  main  = "(B) LDW-CPS1", 
  axes.range = c(-8000, 8000)
)

plot_catt(
  catt1 = catt.ldw$catt,
  catt2 = catt.psid$catt,
  att1  = catt.ldw$att[1],
  att2  = catt.psid$att[1],
  xlab  = "CATT (Experimental)",
  ylab  = "CATT (PSID1)",
  main  = "(C) LDW-PSID1",
  axes.range = c(-8000, 8000)
)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA11. SubfigureD-H:Top-LDW-CPS1. SubfigureI-M:Top-LDW-PSID1. CATT Estimates Model A using LDW Data: Experimental vs. Nonexperimental"}
# plot results
par(mfrow = c(2,2)) 
par(cex.main = 0.9)

plot_catt_panels(
  exp_catt = catt.ldw,
  catt_list  = catt.top5_cps,
  plot_titles = paste0("(", LETTERS[4:8], ") Top CPS1: ", top5_methods.cps)
)

plot_catt_panels(
  exp_catt = catt.ldw,
  catt_list  = catt.top5_psid,
  plot_titles = paste0("(", LETTERS[9:13], ") Top PSID1: ", top5_methods.psid)
)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA11. SubfigureN:LDW-CPS1-PLUS. SubfigureO:LDW-PSID1-PLUS. CATT Estimates Model A using LDW Data: Experimental vs. Nonexperimental"}
# plot results
par(mfrow = c(2,2)) 
par(cex.main = 0.9)

plot_catt(
  catt1 <- catt.trim_cps$catt,
  catt2 <- catt.cps_trim$catt,
  att1 <- catt.trim_cps$att[1],
  att2 <- catt.cps_trim$att[1],
  xlab  = "CATT (Experimental)",
  ylab  = "CATT (CPS1-PLUS-Trimmed)",
  main  = "(N) Trimmed LDW-CPS1-PLUS", 
  axes.range = c(-8000, 8000)
)

plot_catt(
  catt1 <- catt.trim_psid$catt,
  catt2 <- catt.psid_trim$catt,
  att1 <- catt.trim_psid$att[1],
  att2 <- catt.psid_trim$att[1],
  xlab  = "CATT (Experimental)",
  ylab  = "CATT (PSID1-PLUS-Trimmed)",
  main  = "(O) Trimmed LDW-PSID1-PLUS",
  axes.range = c(-8000, 8000)
)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA12. CATT Estimates Model A using LDW Data"}
# combine all catt objects 
all_catt <- c(list(catt.ldw, catt.cps, catt.psid), 
              catt.top5_cps, catt.top5_psid, 
              list(catt.cps_trim, catt.psid_trim))

all_catt_eval <- eval_catt(all_catt, all_plot_titles)
datatable(all_catt_eval, caption = "CATT Summary Statistics",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

Specifically, with LDW-CPS1, CATT estimates span from $-4,979.99 to $7,201.81, contrasting with the CATT estimated from experimental data which ranges from $-189.40	to $3,835.95, with a mean CATT estimate of $1,713.26. The LDW-PSID1 data shows an even broader CATT estimate range, spanning from $-8,643.02 to $4,903.22, and a notably lower mean of approximately $820.77.

Among the trimmed and top-ranked LDW-CPS1 subsamples, CATT ranges vary substantially. Specific samples such as `card` or `overlap_crump` subsamples produce minimum CATTs that are considerably negative, but produce positive mean estimates. Subsamples like `ps_threshold`, and `optimal_pair` also deliver positive mean CATT estimates, though with smaller negative minimums, while `mahvars` is the only method among these to produce a negative mean CATT. Importantly, across all methods, the mean CATT estimates are far away from experimental mean estimate. 

The CATT estimates for the LDW-PSID1 trimmed and top-ranked subsamples reveal substantially decreased mean values and wider extremes compared to their CPS1 counterparts, reflecting greater difficulties in producing reliable treatment effect estimates.

This variation in range and means across methods and samples reflects substantial heterogeneity in treatment effect estimation, indicating that while some criteria improve alignment with experimental benchmarks, others introduce considerable discrepancies and variability in estimated heterogeneous effects.
```{r, message=FALSE, warning=FALSE}
# save results
save_main_catt_panels(
  catt_refs = list(catt.ldw),
  catt_comps = list(catt.cps, catt.psid),
  ylabels = c("CATT (CPS1)", "CATT (PSID1)"),
  prefix = "ldw_model_a_catt_main_panels",
  main_titles = c("(B) LDW-CPS1", "(C) LDW-PSID1")
)

save_catt_panels(
  exp_catt = catt.ldw,
  catt_list  = catt.top5_cps,
  plot_titles = paste0("(", LETTERS[4:8], ") Top CPS1: ", top5_methods.cps),
  prefix = "ldw_model_a_catt_top5_cps"
)

save_catt_panels(
  exp_catt = catt.ldw,
  catt_list  = catt.top5_psid,
  plot_titles = paste0("(", LETTERS[9:13], ") Top PSID1: ", top5_methods.psid),
  prefix = "ldw_model_a_catt_top5_psid"
)

save_plus_catt_panels(
  catt1_list = list(catt.trim_cps, catt.trim_psid),
  catt2_list = list(catt.cps_trim, catt.psid_trim),
  ylabels = c("CATT (CPS1-PLUS-Trimmed)", "CATT (PSID1-PLUS-Trimmed)"),
  prefix = "ldw_model_a_catt_plus_panels",
  main_titles = c("(N) LDW-CPS1-PLUS", "(O) LDW-PSID1-PLUS")
)
```

Although the AIPW estimator can produce ATT estimates closely aligned with the experimental benchmark using LDW data, its performance for revealing the true CATT is considerably worse. 

### Quantile treatment effect on the treated (QTET)

QTET provides a robust way to analyze treatment effect heterogeneity while being less impacted by outliers compared to average treatment effect measures. The QTET estimates are obtained using the propensity score re-weighting method introduced by Firpo (2007). Implementation leverages the `qte()` function from Imbens and Xu (2024), while visualization employs a modified version of their `plot_qte()` function.
```{r, message=FALSE, warning=FALSE}
qte.ldw <- est_qte_safe(Y, treat, covar, data = ldw, cores = 4)
qte.ldw_cps <- est_qte_safe(Y, treat, covar, data = ldw_cps)
qte.ldw_psid <- est_qte_safe(Y, treat, covar, data = ldw_psid)
qte.top5_cps  <- lapply(top5_datasets.cps,  function(d) est_qte_safe(Y, treat, covar, data = d))
qte.top5_psid <- lapply(top5_datasets.psid, function(d) est_qte_safe(Y, treat, covar, data = d))
```

```{r, message=FALSE, warning=FALSE}
qte.trim.ldw_cps <- est_qte_safe(Y, treat, NULL, data = ldw_trim_cps)
qte.trim.ldw_psid <- est_qte_safe(Y, treat, NULL, data = ldw_trim_psid)
qte.ldw_cps.trim <- est_qte_safe(Y, treat, covar, data = ldw_cps_trim) 
qte.ldw_psid.trim <- est_qte_safe(Y, treat, covar, data = ldw_psid_trim) 
```

```{r, message=FALSE, warning=FALSE}
qte.ldw0 <- est_qte_safe(Y, treat, NULL, data = ldw)
qte.ldw_cps0 <- est_qte_safe(Y, treat, NULL, data = ldw_cps)
qte.ldw_psid0 <- est_qte_safe(Y, treat, NULL, data = ldw_psid)
qte.top5_cps0  <- lapply(top5_datasets.cps,  function(d) est_qte_safe(Y, treat, NULL, data = d))
qte.top5_psid0 <- lapply(top5_datasets.psid, function(d) est_qte_safe(Y, treat, NULL, data = d))
```

```{r, message=FALSE, warning=FALSE}
qte.ldw_cps.trim0 <- est_qte_safe(Y, treat, NULL, data = ldw_cps_trim)
qte.ldw_psid.trim0 <- est_qte_safe(Y, treat, NULL, data = ldw_psid_trim)
```

Each plot displays three distinct series for every dataset analyzed: (1) The blue line with diamond-shaped points represents QTET estimates derived from the experimental sample, serving as the benchmark. (2) The red line with triangles shows QTET estimates from the observational data before any adjustment. (3) The black line with circles presents QTET estimates for the observational data after applying covariate adjustment.
```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA13. SubfigureB:LDW-CPS1. SubfigureC:LDW-PSID1. SubfigureD-H:Top-LDW-CPS1. SubfigureI-M:Top-LDW-PSID1. SubfigureN:LDW-CPS1-PLUS. SubfigureO:LDW-PSID1-PLUS. QTET Estimates Model A using LDW Data: Experimental vs. Nonexperimental"}
par(mfrow = c(2,2))
par(cex.main = 0.9)
ylim = c(-25000, 15000)
# CPS1
plot_qte(qte.ldw_cps, qte.ldw_cps0, qte.ldw, main = "(B) LDW-CPS1", ylim)
legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"), 
       lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")

# PSID1
plot_qte(qte.ldw_psid, qte.ldw_psid0, qte.ldw, main = "(C) LDW-PSID1", ylim)
legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"), 
       lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")

# CPS1 top methods
plot_qte_top(qte.top5_cps, qte.top5_cps0, qte.ldw, all_plot_titles, main_start = 4)

# PSID1 top methods
plot_qte_top(qte.top5_psid, qte.top5_psid0, qte.ldw, all_plot_titles, main_start = 9)

# CPS1-PLUS trimmed
plot_qte(qte.ldw_cps.trim, qte.ldw_cps.trim0, qte.trim.ldw_cps, main = "(N) LDW-CPS1-PLUS (Trimmed)", ylim)
legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"), 
       lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")

# PSID1-PLUS trimmed
plot_qte(qte.ldw_psid.trim, qte.ldw_psid.trim0, qte.trim.ldw_psid, main = "(O) LDW-PSID1-PLUS (Trimmed)", ylim)
legend("bottomleft", legend = c("Experimental", "Unadjusted", "Adjusted"), 
    lty = 1, pch = c(16, 17, 16), col = c(4, 2, 1), bty = "n")
```

These figures display QTET estimates derived from both the experimental and various observational samples. The QTETs estimated from the original and trimmed LDW-CPS1 sample (B and D) corresponds well with the true QTET, although the estimates are often underpowered. The QTET estimates from the original and trimmed LDW-PSID1 subsample (C and E) show clear biases when compared to the experimental benchmark, which clusters near zero. Among the top-ranked LDW-CPS1-based subsamples, QTETs (F - J) continue to track the true experimental effect well, whereas LDW-PSID1-based subsamples produce QTETs (K - O) that exhibit increased bias and noticeably wider confidence bands, indicating greater estimation uncertainty.
```{r, message=FALSE, warning=FALSE}
# list results
plots_ldw <- list(
  list(mod = qte.ldw_cps, mod0 = qte.ldw_cps0, bm = qte.ldw, 
       main = "(B) LDW-CPS1"),
  list(mod = qte.ldw_psid, mod0 = qte.ldw_psid0, bm = qte.ldw, 
       main = "(C) LDW-PSID1"),
  list(mod = qte.ldw_cps.trim, mod0 = qte.ldw_cps.trim0, bm = qte.trim.ldw_cps, 
       main = "(N) LDW-CPS1-PLUS (Trimmed)"),
  list(mod = qte.ldw_psid.trim, mod0 = qte.ldw_psid.trim0, bm = qte.trim.ldw_psid, 
       main = "(O) LDW-PSID1-PLUS (Trimmed)")
)

# save results
save_qtet(plots_ldw, prefix = "ldw_model_a")
save_qte_top(qte.top5_cps, qte.top5_cps0, qte.ldw, all_plot_titles, 
             main_start = 8, prefix = "ldw_model_a_top")
save_qte_top(qte.top5_psid, qte.top5_psid0, qte.ldw, all_plot_titles, 
             main_start = 13, prefix = "ldw_model_a_top")
```

### Assessing outcome weights (OW)
```{r, message=FALSE, warning=FALSE}
# list all datasets
all_datasets <- c(list(ldw, ldw_cps, ldw_psid), top5_datasets.cps, top5_datasets.psid, list(ldw_trim_cps, ldw_trim_psid))
```

```{r, message=FALSE, warning=FALSE}
# estimate ATT 
res_att <- get_res_att(all_datasets, Y, treat, covar)

# extract outcome weights
ow_att <- derive_ow(res_att_valid)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA14. Outcome Weights Model A using LDW Data"}
par(mfrow = c(2,2))
par(cex.main = 0.9)

# plot outcome weights distribution
plot_ow(ow_att, all_plot_titles_valid) 
```

```{r, message=FALSE, warning=FALSE}
# evaluate results
res_ow <- eval_ow(ow_att, all_datasets_valid, all_plot_titles_valid, treat, "AIPW-ATT")

# print results
datatable(res_ow, caption = "Outcome Weights for Treated and Untreated",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

<div class="callout-caution">
The outcome weights are not obtained directly from the ATT point estimates due to computational design of other methods. Instead, the ATT is separately estimated for each dataset following the approach developed by @OW_Package.
</div>

The evaluation reveals that across each sample the estimated outcome weights sum to one within the treated group and minus one within the untreated group, resulting in an overall sum of zero.
```{r, message=FALSE, warning=FALSE}
#save results
save_ow(ow_att, all_plot_titles, prefix = "model_a")
```

## Validation through placebo analyses

To further evaluate the credibility of the unconfoundedness assumption, a placebo analysis is performed by designating 1975 earnings (re75) as the outcome variable and omitting both re75 and u75 from the list of covariates. The analysis utilizes all previously considered samples and recomputes the ATT via the function `estimate_all`, conditioning only on the remaining set of covariates.
```{r, message=FALSE, warning=FALSE}
# define variables
Y <- "re75"
treat <- "treat"
covar <- c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "u74")
```

```{r, message=FALSE, warning=FALSE}
# estimate placebo ATT on original and observational datasets
out1_pl <- estimate_all(ldw, Y, "treat", covar)
out2_pl <- estimate_all(ldw_cps, Y, "treat", covar)
out3_pl <- estimate_all(ldw_psid, Y, "treat", covar)
```

```{r, message=FALSE, warning=FALSE}
# estimate placebo ATT on top ranked datasets
out.cps_pl <- lapply(top5_datasets.cps, function(d) estimate_all(d, Y, "treat", covar))
out.psid_pl <- lapply(top5_datasets.psid, function(d) estimate_all(d, Y, "treat", covar))

out4_pl <- out.cps_pl[[1]]
out5_pl <- out.cps_pl[[2]]
out6_pl <- out.cps_pl[[3]]
out7_pl <- out.cps_pl[[4]]
out8_pl <- out.cps_pl[[5]]

out9_pl <- out.psid_pl[[1]]
out10_pl <- out.psid_pl[[2]]
out11_pl <- out.psid_pl[[3]]
out12_pl <- out.psid_pl[[4]]
out13_pl <- out.psid_pl[[5]]
```

```{r, message=FALSE, warning=FALSE}
# estimate placebo ATT on plus datasets
load("data/trimmed.RData")
out14_pl <- estimate_all(ldw_trim_cps_pl, Y, "treat", covar)
out15_pl <- estimate_all(ldw_trim_psid_pl, Y, "treat", covar)
out16_pl <- estimate_all(ldw_cps_trim_pl, Y, "treat", covar)
out17_pl <- estimate_all(ldw_psid_trim_pl, Y, "treat", covar)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA15. SubfigureA:LDW-Experimental. SubfigureB:LDW-CPS1. SubfigureC:LDW-PSID1. SubfigureD-H:Top-LDW-CPS1. SubfigureI-M:Top-LDW-PSID1. Placebo Test Model A: '75 Earnings as the Outcome"}
par(mfrow = c(4, 1), mar = c(4, 4, 2, 1))

# experimental benchmarks
band.exp_pl <- out1_pl[1, 3:4]
est.exp_pl <- out1_pl[1, 1]

# plot placebo results
plot_coef(out1_pl,  band = band.exp_pl, line = est.exp_pl, 
          ylim = c(-15500, 5500), main = "(A) LDW-Experimental")

plot_coef(out2_pl,  band = band.exp_pl, line = est.exp_pl, 
          ylim = c(-15500, 5500), main = "(B) LDW-CPS1")

plot_coef(out3_pl,  band = band.exp_pl, line = est.exp_pl,
          ylim = c(-15500, 5500), main = "(C) LDW-PSID1")

for (i in seq_along(out.cps_pl)) {
  this_title <- paste0("(", LETTERS[i+3], ") Top CPS1: ", top5_methods.cps[i])
  plot_coef(out.cps_pl[[i]], band = band.exp_pl, line = est.exp_pl,
            ylim = c(-15500, 5500), main = this_title)
}

for (i in seq_along(out.psid_pl)) {
  this_title <- paste0("(", LETTERS[i+8], ") Top PSID1: ", top5_methods.psid[i])
  plot_coef(out.psid_pl[[i]], band = band.exp_pl, line = est.exp_pl,
            ylim = c(-15500, 5500), main = this_title)
}
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA15. SubfigureN:LDW-CPS1-PLUS. SubfigureO:LDW-PSID1-PLUS. Placebo Test Model A: '75 Earnings as the Outcome"}
par(mfrow = c(4, 1))

# nonexperimtenal benchmarks
band.cps_plus_pl <- out14_pl[1, 3:4]
est.cps_plus_pl <- out14_pl[1, 1]
band.psid_plus_pl <- out15_pl[1, 3:4]
est.psid_plus_pl <- out15_pl[1, 1]

# plot results
plot_coef(out16_pl, band = band.cps_plus_pl, line = est.cps_plus_pl, 
          ylim = c(-12000, 2000), main = "(N) Trimmed LDW-CPS1-PLUS")

plot_coef(out17_pl, band = band.psid_plus_pl, line = est.psid_plus_pl, 
          ylim = c(-12000, 2000), main = "(O) Trimmed LDW-PSID1-PLUS")
```

```{r, message=FALSE, warning=FALSE}
# save results
save_att_panels(
  out_list = list(out1_pl, out2_pl, out3_pl),
  plot_titles = c("(A) LDW-Experimental", "(B) LDW-CPS1", "(C) LDW-PSID1"),
  band_list = list(band.exp_pl, band.exp_pl, band.exp_pl),
  est_list  = list(est.exp_pl, est.exp_pl, est.exp_pl),
  prefix = "ldw_model_a_pl_est_exp"
)

save_att_panels(
  out_list = out.cps_pl,
  plot_titles = paste0("(", LETTERS[4:8], ") Top CPS1: ", top5_methods.cps),
  band_list = replicate(length(out.cps_pl), band.exp_pl, simplify = FALSE),
  est_list  = replicate(length(out.cps_pl), est.exp_pl, simplify = FALSE),
  prefix = "ldw_model_a_pl_est_top_cps"
)

save_att_panels(
  out_list = out.psid_pl,
  plot_titles = paste0("(", LETTERS[9:13], ") Top PSID1: ", top5_methods.psid),
  band_list = replicate(length(out.psid_pl), band.exp_pl, simplify = FALSE),
  est_list  = replicate(length(out.psid_pl), est.exp_pl, simplify = FALSE),
  prefix = "ldw_model_a_pl_est_top_psid"
)

save_att_panels(
  out_list = list(out16_pl, out17_pl),
  plot_titles = c("(P) Trimmed LDW-CPS1-PLUS", "(Q) Trimmed LDW-PSID1-PLUS"),
  band_list = list(band.cps_plus_pl, band.psid_plus_pl),
  est_list  = list(est.cps_plus_pl, est.psid_plus_pl),
  prefix = "ldw_model_a_pl_est_plus",
  ylim = c(-12000, 2000)
)
```

The placebo ATT results are presented in the table below:
```{r, message=FALSE, warning=FALSE}
# get placebo results
all_outs.pl <- c(list(out1_pl, out2_pl, out3_pl), out.cps_pl, out.psid_pl, list(out16_pl, out17_pl))
result_mat_pl <- create_matrix_results(all_outs.pl, all_plot_titles)

# print results
datatable(result_mat_pl, caption = "Placebo ATT Estimates and SEs",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

The placebo analysis reveals that the experimental benchmarks are positive, near zero and statistically insignificant, while all estimators applied to the observational datasets, LDW-CPS1 and LDW-PSID1, yield by majority large, negative ATT estimates.

For LDW-CPS1, most ATT estimates are negative. The `overlap_crump` subsample shows a modest improvement toward the experimental benchmark compared to LDW-CPS1 (B) results, but also produces notably poorer estimates for certain estimators. The trimmed subsamples (D) and (I), which use similar approaches, further improve upon the LDW-CPS1 (B) results. The `optimal_pair` and `card` subsamples provide even greater improvement. The `mahvars` (G) subsample delivers ATT estimates closest to the experimental benchmark and is the only case to yield positive ATT estimates across the LDW-CPS1 samples.

For LDW-PSID1, the subsample (O) applying the `overlap_common_range` method shows no improvement compared to the LDW-PSID1 sample (C) except for the `Diff-in-Means` estimator. The `overlap_crump` subsample generally delivers ATT estimates closer to the experimental benchmark, while the `card` and `mahvars` subsamples producing estimates even closer to the experimental benchmark. The trimmed subsamples (E) and (N), which use similar approaches, achieve the closest alignment with the benchmark.

Across most estimators and observational top-ranked subsamples (F–O) the placebo analysis reveals substantial bias and deviation from the true effect, highlighting the persistent challenges in adjusting for confounding using observational data. This underlines the limitations of these methods in recovering unbiased ATT estimates outside of randomized experimental settings.
```{r, message=FALSE, warning=FALSE}
# save results 
save_csv(result_mat_pl, "ldw_att_estimates_pl_model_a")
```

## Validation through sensitivity analyses

Finally, a sensitivity analyses using the LDW data is conducted and results are depicted in contour plots using the function `sens_ana` of Imbens & Xu (2024).
```{r, message=FALSE, warning=FALSE}
# define variables
Y <- "re78"
treat <- "treat"
covar <- c("age", "education", "black", "hispanic", "married", "nodegree", "re75", "u74", "u75")
bm <- c("re75") 
```

```{r, message=FALSE, warning=FALSE}
# check for valid datasets 
filtered_datasets_sens <- check_filter_datasets(all_datasets, Y, treat, covar, bm)
```

```{r, message=FALSE, warning=FALSE, out.width='100%', fig.asp=1, fig.cap="FIGUREA16. Sensitivity Analyses Model A"}
par(mfrow = c(2,2))
par(cex.main = 0.8)

# loop over valid datasets and assign index
for (i in seq_along(filtered_datasets_sens)) {
    idx <- which(sapply(all_datasets, identical, filtered_datasets_sens[[i]]))
    tryCatch({
      sens_ana(filtered_datasets_sens[[i]], Y, treat, covar, bm, kd = 1:3)
      title(main = all_plot_titles[idx])
    }, error = function(e) {
      cat(sprintf("Skipped %s: %s\n", all_plot_titles[idx], e$message))
      plot.new()
      text(0.5, 0.5, paste("Insufficient sample size\n", all_plot_titles[idx]), cex = 1.2)
    })
}
```

```{r, message=FALSE, warning=FALSE}
# save results
save_sensitivity_plots(filtered_datasets_sens, Y, treat, covar, bm, all_plot_titles, "ldw_model_a")
```

The sensitivity analysis shows that for most samples, the estimated treatment effects are fairly robust to increasing confounder strength, as indicated by relatively stable values despite up to triple the correlation levels of `re75`. Notably, some samples such as LDW-PSID1 exhibit a greater sensitivity with more pronounced swings including negative estimates at higher confounder strengths, contrasting with consistent positive effects in LDW-CPS1 variants. This highlights sample-specific differences in robustness against potential unobserved confounding.

<div class="callout-tip">
The analyses suggest that some samples maintain stable treatment effect estimates despite increasing confounder strength, while others show greater sensitivity. This highlights the importance of assessing robustness across different samples (respectively methods used to derive specific samples) to account for potential unobserved bias.
</div>

## Inspection of re74 and re75
### Correlation 
```{r, message=FALSE, warning=FALSE}
# compute correlation 
cor_results <- sapply(all_datasets, function(d) cor(d$re74, d$re75, use = "complete.obs"))
cor_table <- data.frame(Method = all_plot_titles, Correlation = cor_results)

# print results
datatable(cor_table, caption = "Correlation between re74 and re75 across datasets",
          options = list(scrollX = TRUE, paging = FALSE, pageLength = -1, searching = FALSE))
```

### Distributional balance
```{r, message=FALSE, warning=FALSE, fig.width=12, fig.height=3, out.width='100%'}
# plot re75 distributions for top5 CPS methods
for (i in seq_along(top5_datasets.cps)) {
  tryCatch({
    plot_title <- paste0("(", LETTERS[i+3], ") CPS1: ", top5_methods.cps[i], " - re75")
    print(bal.plot(treat ~ re75, data = top5_datasets.cps[[i]], which = "both") + 
            ggplot2::ggtitle(plot_title))
  }, error = function(e) {
    cat(sprintf("Could not plot %s: %s\n", top5_methods.cps[i], e$message))
  })
}

# plot re75 distributions for top5 PSID methods
for (i in seq_along(top5_datasets.psid)) {
  tryCatch({
    plot_title <- paste0("(", LETTERS[i+8], ") PSID1: ", top5_methods.psid[i], " - re75")
    print(bal.plot(treat ~ re75, data = top5_datasets.psid[[i]], which = "both") + 
            ggplot2::ggtitle(plot_title))
  }, error = function(e) {
    cat(sprintf("Could not plot %s: %s\n", top5_methods.psid[i], e$message))
  })
}
```

```{r, message=FALSE, warning=FALSE, fig.width=12, fig.height=3, out.width='100%'}
# plot re74 distributions for top5 CPS methods
for (i in seq_along(top5_datasets.cps)) {
  tryCatch({
    plot_title <- paste0("(", LETTERS[i+3], ") CPS1: ", top5_methods.cps[i], " - re74")
    print(bal.plot(treat ~ re74, data = top5_datasets.cps[[i]], which = "both") + 
            ggplot2::ggtitle(plot_title))
  }, error = function(e) {
    cat(sprintf("Could not plot %s: %s\n", top5_methods.cps[i], e$message))
  })
}

# plot re74 distributions for top5 PSID methods
for (i in seq_along(top5_datasets.psid)) {
  tryCatch({
    plot_title <- paste0("(", LETTERS[i+8], ") PSID1: ", top5_methods.psid[i], " - re74")
    print(bal.plot(treat ~ re74, data = top5_datasets.psid[[i]], which = "both") + 
            ggplot2::ggtitle(plot_title))
  }, error = function(e) {
    cat(sprintf("Could not plot %s: %s\n", top5_methods.psid[i], e$message))
  })
}
```

## Summary
After reexamining model A, that is based on the LaLonde-Dehejia-Wahba (LDW) data and its augmented versions with control groups from CPS-1 and PSID-1, several insights into causal inference challenges emerge. First, we confirm that ensuring overlap between treated and control groups remains crucial for credible causal estimation. The LDW-Experimental data exhibit excellent overlap, while the observational datasets (LDW-CPS1 and LDW-PSID1) show weaker overlap, with many treated units outside the control range. Augmenting these datasets with experimental controls improves overlap but does not consistently improve covariate balance.

Second, controlling for covariates such as age, education, race, and prior earnings, and using various methods including matching, weighting, truncation, trimming, and combined approaches, largely yields consistent effects. The propensity score plays an important role in assessing overlap and balancing covariates across groups.

Third, the LDW dataset is somewhat unique in that most methods approximate the experimental benchmark well for average treatment effects on the treated (ATT), an achievement not fully replicated in the original LaLonde samples. However, placebo tests using pre-treatment earnings and sensitivity analyses reveal lingering concerns about the unconfoundedness assumption and robustness of some estimators.

Overall, this chapter highlights the importance of overlap and covariate balance, the utility of propensity scores, and the need for rigorous validation of treatment assignment assumptions to produce credible causal estimates. Despite improvements in data and methods, tests of unconfoundedness via placebo outcomes suggest caution in interpreting causal effects.