---
title: "Replication and Extension of Imbens & Xu (2024) Tutorial - LaLonde Males Samples" 
output: html_document
date: "2025-05-08"
---

# LaLonde-Calónico-Smith (LCS) Data 
This section (5) examines the LaLonde female samples reconstructed by Calónico and Smith (2017), referred to as the LaLonde-Calónico-Smith (LCS) sample (loaded as `lcs` below, similar to Imbens & Xu). 

For detailed explanations of the analysis steps and tips, please refer to section 2. Here, we only present and explain the lcs–specific results.

# Set up
## Source functions and load data
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# source functions
source("https://github.com/xuyiqing/lalonde/blob/main/tutorial/functions.R?raw=TRUE")
source("functions.R")
```

```{r}
# load data
load("data/lcs.RData")
```

```{r}
# set seed
set.seed(42)
```

## Load and preprocess data 
```{r}
#| code-fold: show
# expc = 0: experimental treated; 
# expc = 1: experimental control; 
# expc = 2: psid control;

lcs_psid$expc <- 0 
lcs_psid[lcs_psid$treat==0, ]$expc <- 2 
lcs_tr <- lcs[lcs$treat==1, ]
lcs_co <- lcs[lcs$treat==0, ]
lcs_co$treat <-  1
lcs_co$expc <- 1
lcs_psid_plus <- rbind.data.frame(lcs_psid, lcs_co)
```

```{r}
#| code-fold: show
# define variables
Y <- "re79"
treat <- "treat"
# redefine covariates: removing "nchildren75" to be used as placebo outcome
covar <- c("age", "educ", "nodegree", "married", "black", "hisp", 
           "re75", "u75")
```

::: callout-tip
In the following analysis, only PSID1 data is used as the original LaLonde study and its reconstruction by Calónico and Smith (2017) designate PSID1 as the appropriate nonexperimental control group for women, providing a comparable observational dataset that aligns with the experimental sample’s characteristics.
:::

## Assessing overlap and covariate balance
### Overlap
```{r, class.source = "fold-show", fig.cap='FIGURE1. SubfigureA:LCS-Experimental. SubfigureB:LCS-PSID1.',out.width='100%', fig.asp=0.5}
#| code-fold: show
# assess overlap
lcs.ps <- assess_overlap(data = lcs, treat = treat, cov = covar, xlim = c(-1.5, 1.5), breaks = 40)
lcs_psid.ps <- assess_overlap(data = lcs_psid, treat = treat, cov = covar, xlim = c(-11, 7), breaks = 40)
```

As anticipated, the LCS-Experimental data exhibit an almost perfect overlap. In contrast, the observational dataset LCS-PSID1 displays weak overlap. 
```{r, class.source = "fold-show", fig.cap='FIGURE2. SubfigureC:LCS-PSID1-PLUS.', out.width='80%', fig.asp=1, fig.align='center'}
#| code-fold: show
# assess overlap
lcs_psid_plus.ps <- assess_overlap(data = lcs_psid_plus, treat = treat, cov = covar, xlim = c(-15, 5))
```

With the expanded dataset LCS-PSID1, it is evident that the degree of overlap between treated and control groups has improved, as seen by a greater spread of log-odds densities across both samples.

```{r}
#| code-fold: show
# set model formula
model <- as.formula(paste(treat, "~", paste(covar, collapse = " + ")))
```

### Covariate balance
```{r, message=FALSE, warning=FALSE}
love.plot(lcs, lcs_psid, treat, covar = covar, title = "LCS-PSID1")
love.plot(lcs, lcs_psid_plus, treat, covar = covar, title = "LCS-PSID1-PLUS")

bal.tab(model, data = lcs_psid, treat = "treat")
bal.tab(model, data = lcs_psid_plus, treat = "treat")
```

ADD EXPLANATION HERE

## Improving primarily covariate balance
### Matching 
#### Distance Matching
##### 1:1 Nearest neighbor matching
```{r}
#| code-fold: show
# perform nearest neighbor matching with k=1, logistic propensity score and replacement
#m.out.psid_plus.nearest <- matchit(model, data = lcs_psid_plus, method = "nearest", distance = "logit", replace = TRUE)
```

##### k:1 matching (k=2)
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
k<-2
# perform nearest neighbor matching with k=2, logistic propensity score and replacement
#m.out.psid_plus.k2 <- matchit(model, data = lcs_psid_plus, method = "nearest", distance = "logit", ratio = k, replace = TRUE)
```

##### k:1 matching (k=3)
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
k<-3
# perform nearest neighbor matching with k=3, logistic propensity score and replacement
#m.out.psid_plus.k3 <- matchit(model, data = lcs_psid_plus, method = "nearest", distance = "logit", ratio = k, replace = TRUE)
```

##### Caliper matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform nearest neighbor matching with a caliper of 0.1 on the logistic propensity score 
#m.out.psid_plus.caliper <- matchit(model, data = lcs_psid_plus, method = "nearest", distance = "logit", caliper = 0.1, replace = TRUE)
```

##### Common support restriction matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform nearest neighbor matching with exclusion of units outside common support
#m.out.psid_plus.cs <- matchit(model, data = lcs_psid_plus, method = "nearest", distance = "logit", discard = "both", replace = TRUE)
```

##### Mahalanobis distance matching (mahvars) 
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform nearest neighbor matching using mahalanobis distance on selected covariates
#m.out.psid_plus.mahvars <- matchit(model, data = lcs_psid_plus, method = "nearest", distance = "logit", caliper = 0.2, mahvars = ~ age + educ + re75, replace = FALSE)
```

##### Optimal pair matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform optimal pair matching that minimizes total within-pair distance on propensity scores
m.out.psid_plus.optimal_pair <- matchit(model, data = lcs_psid_plus, method = "optimal", distance = "logit")
```

##### Optimal full matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform optimal full matching allowing sets with varying ratios of treated to controls and minimizing a global distance criterion
m.out.psid_plus.optimal_full <- matchit(model, data = lcs_psid_plus, method = "full", distance = "logit")
```

##### Generalized full matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform generalized full matching using fast approximation and producing matched sets with flexible treated/control ratios 
m.out.psid_plus.general_full <- matchit(model, data = lcs_psid_plus, method = "quick", distance = "logit")
```

##### Genetic matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# perform genetic matching 
#m.out.psid_plus.genetic <- matchit(model, data = lcs_psid_plus, method = "genetic", distance = "logit", replace = TRUE, pop.size = 100)
```

#### Stratum matching
##### Exact matching (exact)
Strata = unique covariate profiles (raw covariates)
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# match units exactly by raw covariate profiles 
#m.out.psid_plus.exact <- matchit(model, data = lcs_psid_plus, method = "exact")
```

##### Coarsened matching (cem)
Strata = coarsened versions of covariates
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# match units exactly within coarse strata 
m.out.psid_plus.cem <- matchit(model, data = lcs_psid_plus, method = "cem")
```

##### Subclassification 
Strata = bins of the propensity score
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# partition sample into fixed number of bins based on propensity score 
m.out.psid_plus.subcl <- matchit(model, data = lcs_psid_plus, method = "subclass", subclass = 5)
```

#### Pure subset selection
##### Cardinality profile matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# select largest balanced subsample meeting covariate balance tolerances
m.out.psid_plus.card <- matchit(model, data = lcs_psid_plus, method = "cardinality", tols = 0.1, ratio = 1)
```

##### Profile matching
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# matching by directly optimizing balance profile measures across covariates
m.out.psid_plus.profile <- matchit_profile(lcs_psid_plus, treat, covar)
```

### Weighting
#### Inverse probability weights (IPW)
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute weights as inverse of estimated propensity scores 
w.out.psid_plus.ipw <- weightit(model, data = lcs_psid_plus, method = "ps", estimand = "ATT")

lcs_psid_plus$ipw_weight <- w.out.psid_plus.ipw$weights
```

#### Standardized mortality ratio (SMR) treated weights
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# calculate weights for ATT by assigning weight = 1 to treated units and weights proportional to odds of treatment for control
lcs_psid_plus$smr_weight <- create_smr_weights(lcs_psid_plus, model, "ATT")
```

#### Matching weights
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# derive optimal matching weights intending to minimize covariate imbalance while targeting ATT
w.out.psid_plus.opt <- weightit(model, data = lcs_psid_plus, method = "optweight", estimand = "ATT")

lcs_psid_plus$opt_weight <- w.out.psid_plus.opt$weights
```

#### Overlap weights
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# calculate overlap weights emphasizing units with propensity scores near 0.5
lcs_psid_plus$overlap_weight <- create_overlap_weights(lcs_psid_plus, model)
```

#### Entropy weights
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute entropy balancing weights 
w.out.psid_plus.ebal <- weightit(model, data = lcs_psid_plus, method = "ebal", estimand = "ATT")

lcs_psid_plus$ebal_weight <- w.out.psid_plus.ebal$weights
```

## Improving primarily overlap
### Truncation
```{r}
#| code-fold: show
# list of weight columns to apply truncation 
weight_columns <- c("ipw_weight", "smr_weight", "opt_weight", "overlap_weight", "ebal_weight")
```

#### Fixed maximum value truncation
```{r}
#| code-fold: show
# truncate weights by imposing a maximum threshold of 10
lcs_psid_plus.fixed <- lcs_psid_plus
for (wcol in weight_columns) {
  if (wcol %in% names(lcs_psid_plus.fixed)) {
    lcs_psid_plus.fixed <- truncate_weights_fixed(lcs_psid_plus.fixed, weight_col = wcol, max_weight = 10)
  }
}
```

#### At percentile truncation
```{r}
#| code-fold: show
# truncate weights such that values above the 99th percentile are capped
lcs_psid_plus.percent <- lcs_psid_plus
for (wcol in weight_columns) {
  if (wcol %in% names(lcs_psid_plus.percent)) {
    lcs_psid_plus.percent <- truncate_weights_percentile(lcs_psid_plus.percent, weight_col = wcol, percentile = 0.99)
  }
}
```

#### Adaptive weight truncation
```{r}
#| code-fold: sh
# inspect variance of weights 
for (wcol in weight_columns) {
  if (wcol %in% names(lcs_psid_plus)) {
    check_weights(lcs_psid_plus, wcol)
  } else {
    cat("Column", wcol, "not found in lcs_cps.plus\n")
  }
}
```

Regarding these results we can apply adaptive weight truncation to all weights.
```{r}
#| code-fold: show
# truncate adaptively at mean + 3 standard deviations 
lcs_psid_plus.adapt <- lcs_psid_plus
for (wcol in weight_columns) {
  if (wcol %in% names(lcs_psid_plus.adapt)) {
    lcs_psid_plus.adapt <- truncate_weights_adaptive(lcs_psid_plus.adapt, weight_col = wcol, c = 3)
  }
}
```

### Trimming 
#### Propensity score threshold trimming (analogous to tutorial of Imbens & Xu (2024))
```{r}
#| code-fold: show
# apply trimming with threshold 0.9
lcs_psid_trim <- ps_trim(lcs_psid_plus.ps, threshold = 0.9)

# exclude experimental controls, subset trimmed data appropriately
lcs_psid.trim_match <- subset(lcs_psid_trim, expc %in% c(0,2) & ps_assoverlap)

# re-estimate propensity scores on trimmed data and perform 1:1 matching
lcs_psid.trim_match <- psmatch(data = lcs_psid.trim_match, Y = "re79", treat = "treat", cov = covar)

# trim experimental data and re-assign treat variable for controls in sample 3 or 4 (non-treated group)
#lcs_trim_psid <- subset(lcs_psid_trim, expc %in% c(0, 1))
#lcs_trim_psid$treat[lcs_trim_psid$expc == 1] <- 0
```

#### Common range trimming
```{r}
#| code-fold: show
# trim observations outside the common support region of propensity scores
lcs_psid_plus.common  <- common_range_trim(lcs_psid_plus.ps)
```

#### Propensity score trimming (Crump)
```{r}
#| code-fold: show
# trim observations with propensity scores outside [0.1, 0.9] interval
lcs_psid_plus.crump <- crump_trim(lcs_psid_plus.ps, lower = 0.1, upper = 0.9)
```

#### Stuermer trimming
```{r}
#| code-fold: show
# trim observations based on propensity score quantiles separately for treated and control
lcs_psid_plus.stuermer <- stuermer_trim(lcs_psid_plus.ps)
```

#### Walker trimming
```{r}
#| code-fold: show
# trim observations based on preference scores that adjust for treatment prevalence using logit transformations
lcs_psid_plus.walker  <- walker_trim(lcs_psid_plus.ps)
```

### Combination of methods
```{r}
#| code-fold: show
# list trimming methods
trim_names <- c("ps_threshold", "common_range", "stuermer", "walker", "crump")
trimmed_psid <- list(ps_threshold = lcs_psid_trim, common_range = lcs_psid_plus.common, stuermer = lcs_psid_plus.stuermer, walker = lcs_psid_plus.walker, crump = lcs_psid_plus.crump)
```

#### IPW with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump
```{r}
# apply inverse probability weighting (IPW) with trimming and attach IPW weights
ipw_comb.psid_plus <- setNames(
  lapply(trim_names, function(nm) trim_attach_weights(trimmed_psid[[nm]], lcs_psid_plus, "ipw_weight")),
  trim_names
)
```

#### SMR treated weights with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump
```{r}
#| code-fold: show
# apply standardized mortality ratio (SMR) weighting with trimming and attach SMR weights
smr_treat_comb.psid_plus <- setNames(
  lapply(trim_names, function(nm) trim_attach_weights(trimmed_psid[[nm]], lcs_psid_plus, "smr_weight")),
  trim_names
)
```

#### Overlap weights with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump
```{r}
#| code-fold: show
# apply overlap weighting with trimming and attach overlap weights
ov_comb.psid_plus <- setNames(
  lapply(trim_names, function(nm) trim_attach_weights(trimmed_psid[[nm]], lcs_psid_plus, "overlap_weight")),
  trim_names
)
```

#### Entropy weights with 1) propensity threshold 2) common range 3) stuermer 4) walker 5) crump
```{r}
#| code-fold: show
# apply entropy balancing weights with trimming and attach entropy weights
entropy_comb.psid_plus <- setNames(
  lapply(trim_names, function(nm) trim_attach_weights(trimmed_psid[[nm]], lcs_psid_plus, "ebal_weight")),
  trim_names
)
```

## Reassessing methods
### Matching
```{r}
#| code-fold: show
# list all matching methods
methods.psid_plus <- list(
  #nn = m.out.psid_plus.nearest,
  #k2 = m.out.psid_plus.k2,
  #k3 = m.out.psid_plus.k3,
  #caliper = m.out.psid_plus.caliper,
  #cs = m.out.psid_plus.cs,
  #mahvars = m.out.psid_plus.mahvars,
  optimal_pair = m.out.psid_plus.optimal_pair,
  optimal_full = m.out.psid_plus.optimal_full,
  gen_full = m.out.psid_plus.general_full,
  #genetic = m.out.psid_plus.genetic,
  #exact = m.out.psid_plus.exact,
  cem = m.out.psid_plus.cem,
  card = m.out.psid_plus.card,
  profile = m.out.psid_plus.profile,
  subcl = m.out.psid_plus.subcl
)
```

```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# calculate balance statistics 
bal.psid_plus <- cobalt::bal.tab(model, data = lcs_psid_plus, un = TRUE, weights = methods.psid_plus, s.d.denom = "treated")
```

#### SMD
```{r}
#| code-fold: show
# compute absolute SMD
smd_matchit.psid_plus <- compute_abs_smd_matchit(methods.psid_plus)
```

#### Effective Sample Sizes 
```{r}
#| code-fold: show
# compute effective sample sizes (ESS)
ess_matchit.psid_plus <- compute_ess_matchit(bal.psid_plus)
```

#### Visuals
```{r, class.source = "fold-show", fig.cap='FIGURE3.'}
#| code-fold: show
# visualize covariate balance 
plot_matchit(methods.psid_plus, "LCS-PSID1-PLUS")
```

### Weighting
```{r}
#| code-fold: show
# list all weights
weight.psid_plus <- list(
  ipw = lcs_psid_plus$ipw_weight,
  smr_tr = lcs_psid_plus$smr_weight,
  mw = lcs_psid_plus$opt_weight,
  ow = lcs_psid_plus$overlap_weight,
  ew = lcs_psid_plus$ebal_weight
)
```

#### SMD
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute absolute SMD 
smd_weight.psid_plus <- compute_abs_smd_weight(lcs_psid_plus, "treat", covar, weight_columns)
```

#### ESS
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute ESS
ess_weight.psid_plus <- compute_ess_weight(lcs_psid_plus, "treat", covar, weight_columns)
```

#### Visuals
```{r, message=FALSE, warning=FALSE, class.source = "fold-show", fig.cap='FIGURE4.'}
#| code-fold: show
# visualize covariate balance
plot_weighting_methods(lcs_psid_plus, "treat", covar, weight.psid_plus, "LCS-PSID1-PLUS")
```

### Truncation
Methods
```{r}
#| code-fold: show
# list truncation methods
trunc.psid_plus <- list(
  fix_max_value_trunc.psid_plus = lcs_psid_plus.fixed,
  at_perc_trunc.psid_plus = lcs_psid_plus.percent,
  adap_weight_trunc.psid_plus = lcs_psid_plus.adapt
)
```

#### SMD
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute absolute SMD
smd_trunc.psid_plus <- compute_abs_smd_trunc(trunc.psid_plus, "treat", covar, weight_columns)
```

#### ESS
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute ESS
ess_trunc.psid_plus <- compute_ess_trunc(trunc.psid_plus, "treat", covar, weight_columns)
```

#### Visuals
```{r, message=FALSE, warning=FALSE, class.source = "fold-show", fig.cap='FIGURE5.'}
#| code-fold: show
# visualize covariate balance
plot_trunc_methods(trunc.psid_plus, "treat", covar, weight_columns, "LCS-PSID-PLUS")
```

### Trimming
Methods
```{r}
#| code-fold: show
# list trimming object
trim.psid_plus <- list(
  original = lcs_psid_plus,
  ps_threshold = lcs_psid.trim_match,
  common_range = lcs_psid_plus.common,
  crump = lcs_psid_plus.crump,
  stuermer = lcs_psid_plus.stuermer,
  walker = lcs_psid_plus.walker
)
```

#### SMD
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute absolute SMD
smd_trim.psid_plus <- compute_abs_smd_trim(trim.psid_plus, "treat", covar)
```

#### ESS
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# compute ESS
ess_trim.psid_plus <- compute_ess_trim(trim.psid_plus, "treat", covar)
```

#### Visuals
```{r, message=FALSE, warning=FALSE, class.source = "fold-show", fig.cap='FIGURE6.'}
#| code-fold: show
# visualize overlap
plot_trim(trim.psid_plus, treat, covar)
```

```{r, class.source = "fold-show", fig.cap='FIGURE7.'}
#| code-fold: show
# visualize covariate balance
love.plot(lcs_psid, lcs_psid.trim_match, treat, covar = covar, title = "LCS-PSID1-PLUS - propensity threshold trimming")
love.plot(lcs_psid, lcs_psid_plus.common, treat, covar, title = "LCS-PSID1-PLUS - common range trimming")
love.plot(lcs_psid, lcs_psid_plus.crump, treat, covar, title = "LCS-PSID1-PLUS - crump trimming")
love.plot(lcs_psid, lcs_psid_plus.stuermer, treat, covar, title = "LCS-PSID1-PLUS - stuermer trimming")
love.plot(lcs_psid, lcs_psid_plus.walker,  treat, covar, title = "LCS-PSID1-PLUS - walker trimming")
```

## Integrated methods
```{r}
# list all combined method results
comb_meth.psid_plus <- list(
  ipw = ipw_comb.psid_plus,
  smr_treated = smr_treat_comb.psid_plus,
  overlap = ov_comb.psid_plus,
  entropy = entropy_comb.psid_plus
)
```

#### SMD
```{r}
# compute absolute SMD
smd_all_comb_meth.psid_plus <- compute_smd_all_datasets(comb_meth.psid_plus, "treat", covar)
```

#### ESS
```{r}
# compute ESS
ess_all_comb_meth.psid_plus <- compute_ess_all_datasets(comb_meth.psid_plus, "treat", covar)
```

#### Visuals
```{r, message=FALSE, warning=FALSE, class.source = "fold-show", fig.cap='FIGURE9.'} 
# visualize covariate balance
#plot_comb_love_plots(comb_meth.psid_plus, "treat", covar)
```

ADD EXPLANATION HERE -----

### Top methods and datasets
```{r, message=FALSE, warning=FALSE}
# combine all results
all_psid_plus <- combine_results("psid_plus") 
```

```{r, message=FALSE, warning=FALSE}
# save results 
save_csv(all_psid_plus, "PSID1_PLUS_all_results_lcs")
```

```{r}
ranked_psid_plus <- assess_methods(all_psid_plus)
```

```{r}
# get top 5 methods 
top5_methods.psid_plus <- get_top_methods(ranked_psid_plus, top_n = 5)

# save results 
save_csv(top5_methods.psid_plus, "PSID1_PLUS_top5_methods_lcs")
```

The table shows that CPS1 and PSID1 share the same top five methods, with the exception of one differing method and a variation in their ordering for PSID1.
```{r}
dataset_list_psid <- list(
  "All" = lcs_psid_plus, 
  "original" = lcs_psid_plus, 
  #"nn" = m.out.psid_plus.nearest, 
  #"caliper"= m.out.psid_plus.caliper,
  "card" = m.out.psid_plus.card,
  #"cem" = m.out.psid_plus.cem,
  #"cS" = m.out.psid_plus.cs,
  #"k2" = m.out.psid_plus.k2,
  #"k3" = m.out.psid_plus.k3,
  #"mahvars" = m.out.psid_plus.mahvars,
  "optimal_full" = m.out.psid_plus.optimal_full,
  "optimal_pair" = m.out.psid_plus.optimal_pair,
  "gen_full" = m.out.psid_plus.general_full,
  #"genetic" = m.out.psid_plus.genetic,
  #"exact" = m.out.psid_plus.exact,
  "subcl" = m.out.psid_plus.subcl,
  "profile" = m.out.psid_plus.profile,
  "ipw_weight" = lcs_psid_plus$ipw_weight, 
  "smr_weight" = lcs_psid_plus$smr_weight,
  "opt_weight" = lcs_psid_plus$opt_weight,
  "overlap_weight" = lcs_psid_plus$overlap_weight,
  "ebal_weight" = lcs_psid_plus$ebal_weight,
  "fix_max_value_trunc_ebal_weight" = lcs_psid_plus.fixed,
  "fix_max_value_trunc_ipw_weight" = lcs_psid_plus.fixed,
  "fix_max_value_trunc_opt_weight" = lcs_psid_plus.fixed,
  "fix_max_value_trunc_overlap_weight" = lcs_psid_plus.fixed,
  "fix_max_value_trunc_smr_weight" = lcs_psid_plus.fixed,
  "at_perc_trunc_ebal_weight" = lcs_psid_plus.percent,
  "at_perc_trunc_ipw_weight" = lcs_psid_plus.percent,
  "at_perc_trunc_opt_weight" = lcs_psid_plus.percent,
  "at_perc_trunc_overlap_weight" = lcs_psid_plus.percent,
  "at_perc_trunc_smr_weight" = lcs_psid_plus.percent,
  "adap_weight_trunc_ebal_weight" = lcs_psid_plus.adapt,
  "adap_weight_trunc_ipw_weight" = lcs_psid_plus.adapt,
  "adap_weight_trunc_opt_weight" = lcs_psid_plus.adapt,
  "adap_weight_trunc_overlap_weight" = lcs_psid_plus.adapt,
  "adap_weight_trunc_smr_weight" = lcs_psid_plus.adapt,
  "ps_threshold" = lcs_psid.trim_match,
  "common_range" = lcs_psid_plus.common,
  "stuermer" = lcs_psid_plus.stuermer,
  "walker" = lcs_psid_plus.walker,
  "crump" = lcs_psid_plus.crump,
  "ipw_common_range" = ipw_comb.psid_plus[[1]],
  "ipw_crump"= ipw_comb.psid_plus[[2]],
  "ipw_ps_threshold"= ipw_comb.psid_plus[[3]],
  "ipw_stuermer"= ipw_comb.psid_plus[[4]],
  "ipw_walker" = ipw_comb.psid_plus[[5]],
  "smr_treated_common_range" = smr_treat_comb.psid_plus[[1]],
  "smr_treated_crump"= smr_treat_comb.psid_plus[[2]],
  "smr_treated_ps_threshold" = smr_treat_comb.psid_plus[[3]],
  "smr_treated_stuermer" = smr_treat_comb.psid_plus[[4]],
  "smr_treated_walker" = smr_treat_comb.psid_plus[[5]],
  "overlap_common_range" = ov_comb.psid_plus[[1]],
  "overlap_crump" = ov_comb.psid_plus[[2]],
  "overlap_ps_threshold"  = ov_comb.psid_plus[[3]],
  "overlap_stuermer" = ov_comb.psid_plus[[4]],
  "overlap_walker"= ov_comb.psid_plus[[5]],
  "entropy_common_range" = entropy_comb.psid_plus[[1]],
  "entropy_crump" = entropy_comb.psid_plus[[2]],
  "entropy_ps_threshold" = entropy_comb.psid_plus[[3]],
  "entropy_stuermer" = entropy_comb.psid_plus[[4]],
  "entropy_walker" = entropy_comb.psid_plus[[5]])
```

```{r}
# create the datasets from combined lists for your top 5 methods:
top5_datasets.psid_plus <- create_top5_datasets(dataset_list_psid, top5_methods.psid_plus)
```

```{r}
# save them into .RData files
save_top5_individual_files(dataset_list_psid, top5_methods.psid_plus, prefix = "lcs_psid")
```

### Assessing outcome weights (OW)
```{r, warning=FALSE}
#| code-fold: show
# list all datasets
all_datasets <- c(
  list(lcs, lcs_cps, lcs_psid, lcs_cps.trim_match, lcs_psid.trim_match),
  top5_datasets.cps_plus,
  top5_datasets.psid_plus
)
```

```{r, warning=FALSE}
#| code-fold: show
# estimate ATT 
res_att <- get_res_att(all_datasets, Y, treat, covar)
# extract outcome weights
ow_att <- derive_ow(res_att)
```

```{r, out.width='100%', fig.asp=1, fig.cap="FIGURE A14. Outcome Weights Model A using LCS Data"}
#| code-fold: show
plot_ow(ow_att, plot_titles) 
```

```{r, warning=FALSE}
#| code-fold: show
#save results
eval_ow_prop(ow_att, all_datasets, plot_titles, treat_var = treat, estimand = "AIPW-ATT")
```

ADD EXPLANATION HERE

```{r, warning=FALSE}
#| code-fold: show
#save results
save_ow(ow_att, plot_titles, prefix = "lcs")
```

## Estimating
### Average treatment effect on the treated (ATT)
```{r, message=FALSE, warning=FALSE}
out1 <- estimate_all(lcs, "re79", "treat", covar)
out2 <- estimate_all(lcs_psid, "re79", "treat", covar)
```

```{r, message=FALSE, warning=FALSE}
out3 <- estimate_all(lcs_psid.trim_match, "re79", "treat", covar)
```

```{r, message=FALSE, warning=FALSE}
out.psid_plus <- lapply(top5_datasets.psid_plus, function(d) estimate_all(d, "re79", "treat", covar))

out4 <- out.psid_plus[[1]]
out5 <- out.psid_plus[[2]]
out6 <- out.psid_plus[[3]]
out7 <- out.psid_plus[[4]]
out8 <- out.psid_plus[[5]]
```

```{r}
# build plot titles 
base_titles <- c("(A) LCS-Experimental", "(B) LCS-PSID1" , "(C) Trimmed LCS-PSID1")
top_start <- 4 # D is 4th letter
num_psid <- length(top5_methods.psid_plus)
top_letters_psid <- LETTERS[top_start:(top_start + num_psid - 1)]
top5_titles.psid_plus <- paste0("(", top_letters_psid, ") Top PSID1: ", top5_methods.psid_plus)
plot_titles <- c(base_titles, top5_titles.psid_plus)

# combine all results
all_outs <- c(list(out1, out2, out3), out.psid_plus)
```

```{r, out.width='100%', fig.asp=1, fig.cap="FIGURE 10. ATT Estimates Given Unconfoundedness using LCS Samples"}
# plot results
band <- out1[1, 3:4]
est  <- out1[1, 1]
plot_att_panels(all_outs, plot_titles, band, est)
```

```{r}
# save results
save_att_panels(all_outs, plot_titles, band, est, "lcs")
```

ADD EXPLANATION HERE ----

```{r, warning=FALSE}
# prepare all results in the order shown in the plots
all_summaries <- lapply(all_outs, eval_att)
att_summary <- do.call(rbind, all_summaries)
rownames(att_summary) <- plot_titles
att_summary
```

The ATT results are presented in the table below:
```{r class.source = 'fold-hide'}
#| code-fold: show
# get result matrix
result_mat <- create_matrix_results(all_outs, plot_titles)
knitr::kable(result_mat, booktabs = TRUE, caption = "LCS ATT Estimates and SEs for All Samples and Estimators")
```

ADD EXPLANATION HERE ----

```{r}
# save results 
save_csv(result_mat, "LCS_att_estimates")
```

## Alternative Estimands
### Conditional Average Treatment Effect on the Treated (CATT)
```{r, warning=FALSE}
#| code-fold: show
catt.lcs <- catt(lcs, Y, treat, covar)
catt.psid <- catt(lcs_psid, Y, treat, covar)
catt.psid.trim <- catt(lcs_psid.trim_match, Y, treat, covar)
```

```{r, warning=FALSE}
#| code-fold: show
catt.top5_psid_plus <- lapply(top5_datasets.psid_plus, function(d) catt(d, Y, treat, covar))
```

```{r, out.width='100%', fig.asp=1, fig.cap="FIGURE 11. CATT Estimates using LCS Data"}
#| code-fold: show
# combine all catt objects 
all_catt <- c(list(catt.lcs,catt.psid, catt.psid.trim), catt.top5_psid_plus)

# plot results
plot_catt_panels(all_catt, plot_titles)
```

```{r}
# save results
save_catt_panels(all_catt, plot_titles, prefix = "lcs")
```

ADD EXPLANATION HERE ----

```{r, warning=FALSE}
#| code-fold: show
eval_catt(all_catt, plot_titles)
```

ADD EXPLANATION HERE

### Quantile Treatment Effect on the Treated (QTET)
```{r, message=FALSE, warning=FALSE}
#| code-fold: show
qte.lcs <- est_qte(Y, treat, covar, data = lcs, cores = 4)
```

```{r, message=FALSE, warning=FALSE}
#| code-fold: show
qte.lcs_cps <- est_qte(Y, treat, covar, data = lcs_cps)
qte.lcs_psid <- est_qte(Y, treat, covar, data = lcs_psid)
qte.lcs_psid.trim <- est_qte(Y, treat, covar, data = lcs_psid.trim_match) 
```

```{r, warning=FALSE}
#| code-fold: show
qte.top5_psid_plus <- lapply(top5_datasets.psid_plus, function(d) est_qte(Y, treat, covar, data = d))
```

```{r, warning=FALSE}
#| code-fold: show
qte.lcs0 <- est_qte(Y, treat, NULL, data = lcs)
qte.lcs.psid0 <- est_qte(Y, treat, NULL, data = lcs_psid)
qte.lcs_psid.trim0 <- est_qte(Y, treat, NULL, data = lcs_psid.trim_match)
qte.top5_psid_plus0 <- lapply(top5_datasets.psid_plus, function(d) est_qte(Y, treat, NULL, data = d))
```

```{r, out.width='100%', fig.asp=1, fig.cap="FIGURE 12. QTET Estimates using LCS Data: Experimental vs. Nonexperimental"}
#| code-fold: show
# combine all qte objects in order 
all_qtet <- c(list(qte.lcs, qte.lcs_psid, qte.lcs_psid.trim), qte.top5.cps, qte.top5.psid)

# plot results
plot_qtet_panels(all_qtet, plot_titles, experimental_qte = qte.lcs, plots_per_page = 4, ylim = c(-25000, 15000))

# save results
save_qtet_panels(all_qtet, plot_titles, qte.lcs, plots_per_page = 4, ylim = c(-25000, 15000), prefix = "model_b_qtet")
```

ADD EXPLANATION HERE ----

## Validation through Placebo Analyses
```{r, warning=FALSE}
# define variables
Y_pl <- "nchildren75"
treat <- "treat"
covar_pl <- c("age", "educ", "nodegree", "married", "black", "hisp", "re75", "u75")
```

```{r, warning=FALSE}
#| code-fold: show
set.seed(1234)
load("data/lcs.RData") 
# estimate placebo ATT on original and observational datasets
out1_pl <- estimate_all(lcs, Y_pl, "treat", covar_pl)
out2_pl <- estimate_all(lcs_psid, Y_pl, "treat", covar_pl)
```

```{r, warning=FALSE}
#| code-fold: show
# estimate placebo ATT on trimmed datasets (analogous to tutorial of Imbens & Xu (2024))
out3_pl <- estimate_all(lcs_psid_trim, Y_pl, "treat", covar_pl)
```

```{r, message=FALSE, warning=FALSE}
#| code-fold: show
# estimate placebo ATT on top ranked datasets
out.psid_pl <- lapply(top5_datasets.psid_plus, function(d) estimate_all(d, Y_pl, "treat", covar_pl))
out4_pl <- out.psid_pl[[1]]
out5_pl <- out.psid_pl[[2]]
out6_pl <- out.psid_pl[[3]]
out7_pl <- out.psid_pl[[4]]
out8_pl <- out.psid_pl[[5]]
```

```{r}
#| code-fold: show
# collect all placebo results 
all_outs.pl <- c(list(out1_pl, out2_pl, out3_pl), out.psid_pl)
```

```{r, out.width='100%', fig.asp=1, fig.cap="FIGURE 13. Placebo Test LCS: Number of Children in 1975 as the Outcome"}
#| code-fold: show
# plot placebo results
band_pl <- out1_pl[1, 3:4]
est_pl  <- out1_pl[1, 1]
ylim_pl = c(-1.5, 1)
plot_att_panels(all_outs.pl, plot_titles, band_pl, est_pl, ylim_pl)
```

```{r}
# save results
save_att_panels(all_outs.pl, plot_titles, band_pl, est_pl, "lcs_placebo")
```

```{r class.source = 'fold-hide'}
#| code-fold: show
result_mat_pl <- create_matrix_results(all_outs.pl, plot_titles)
knitr::kable(result_mat_pl, booktabs = TRUE, caption = "LCS Placebo ATT Estimates and SEs for All Samples and Estimators")
```

The placebo analysis reveals that the 
ADD EXPLANATION HERE ----

```{r}
#| code-fold: show
# save results 
save_csv(result_mat_pl, "LCS_att_estimates_pl")
```

## Sensitivity Analyses
```{r, warning=FALSE}
#| code-fold: show
# define variables
Y <- "re79"
treat <- "treat"
# redefine covariates
covar <- c("age", "educ", "nodegree", "married", "black", "hisp", "re75", "u75")
bm <- c("re75")
```

```{r, warning=FALSE}
#| code-fold: show
# check for valid datasets 
datasets_sens <- c(list(lcs, lcs_psid), top5_datasets.psid_plus)
filtered_datasets_sens <- check_filter_datasets(datasets_sens, Y, treat, covar, bm)
```

```{r, out.width='100%', fig.asp=1, fig.cap="FIGURE 14. Sensitivity Analyses LCS"}
#| code-fold: show
par(mfrow = c(1,2))
# loop over valid datasets and assign index
for (i in seq_along(filtered_datasets_sens)) {
    idx <- which(sapply(datasets_sens, identical, filtered_datasets_sens[[i]]))
    sens_ana(filtered_datasets_sens[[i]], Y, treat, covar, bm, kd = 1:3)
    title(main = plot_titles[idx])
}
```

```{r}
# save results
save_sensitivity_plots(filtered_datasets_sens, Y, treat, covar, bm, plot_titles, "lcs")
```

ADD EXPLANATION HERE ----

# Summary
